### 2021-12-01 deliverables/paper.Rmd
Added bare templates for paper and presentation

---
title: "Time Series Bootstrapping For Solar Current"
author: "Evan Kessler, Adam Kiehl, and Boston Lee"
date: "2021/12/16"
output: pdf_document
---

```{r reprod}
set.seed(400)
```

```{r libs}
library(tidyverse)
```

### 2021-12-01 deliverables/presentation.Rmd
Added bare templates for paper and presentation

---
title: "Time Series Bootstrapping For Solar Current"
author: "Evan Kessler, Adam Kiehl, and Boston Lee"
date: "2021/12/9"
output: 
  beamer_presentation: default
---

# Background

### 2021-12-01 data_preprocessing/astronomical_adjustment.py
Fixed bugs and unneeded imports in python

@@ -3,7 +3,6 @@ import math
 
 import pandas as pd
 import numpy as np
-import matplotlib.pyplot as plt
 
 
 class AstronomicalAdjustment:

### 2021-12-01 data_preprocessing/main.py

@@ -46,7 +46,7 @@ def main():
         axis=1,
     )
 
-    pd.to_csv(PROCESSED_CSV)
+    df.to_csv(PROCESSED_CSV)
 
 if __name__ == "__main__":
     main()

### 2021-12-02 data_scraping/data_scraping.py
Changed data to local time, modified script write location

@@ -100,7 +100,7 @@ def main():
     # LA coordinates
     lat, lon = 34.0522, -118.2437
     data = pull_nrel_data_multiple_years(lat, lon, "config.toml")
-    data.to_csv("../data/irradiance_full.csv", index=False)
+    data.to_csv("../data/raw/irradiance_full_raw.csv", index=False)
 
 
 if __name__ == "__main__":

### 2021-12-05 deliverables/paper.Rmd
Added Bootstrapping
Hear I wrote the code for bootstrapping using blocks of length 4 and 12 based on average peak hours. I cannot figure out how to add the data from the python scripts, I don't use python, studio together. Everything should be set up though where I can quickly change the data calls once I can access it. I've also added some comments depicting some other approaches I can take.

@@
 ```
+
+```{r functions}
+np <- function(x, l, b) {
+  blocks <- matrix(NA, nrow = b, ncol = l)
+  for(i in 1:length(x)) {
+    blocks[i,] <- x[(i-1)*l+1:(i*l)]
+  }
+  blocks
+}
+
+nbb <- function(x, l) {
+  B <- 10000
+  b <- floor(length(x)/l)
+  theta_hat_star_nbb <- rep(NA, B)
+  nb_blocks <- np(x, l, b)
+  for(i in 1:B) {
+    # sample blocks
+    idx <- sample(1:b, b, replace = TRUE)
+    theta_hat_star_nbb[i] <- mean(x[idx, ])
+  }
+  ggplot() +
+    geom_histogram(aes(theta_hat_star_nbb))
+
+  st.d <- sd(theta_hat_star_nbb)
+  bias <- mean(theta_hat_star_nbb) - mean(x)
+
+  print(paste("The standard deviation with ", l, " blocks is: ", st.d, ". \n The bias is: ",
+              bias))
+}
+
+```
+
+```{r sampling}
+#B1 <- df[df$Hour == c(7:10),]$current_output # Hours before peak
+#B2 <- df[df$Hour == c(11:14),]$current_output # Peak hours
+#B3 <- df[df$Hour == c(15:18),]$current_output # Hours after peak
+
+hours <- df[df$Hour == c(7:18),]$current_output
+nbb(hours, 4) # in order to make b blocks I would need to solve for b = [n/l] (floor) where n is the length of our data and l is 4, so make nbb call l instead of b
+
+nbb(hours, 12)
+
+# Could also block instead by hours but by total current output.
+Peak hours would be a frame around the highest output and then hours before and after would coincide with that.
+This would cause the hours to potentially be different based on the season/month as months that get dark earlier could see peak hours earlier in the day.
+
+#dpeak <- c(B1,B2,B3)
+```

### 2021-12-05 deliverables/presentation.Rmd
Initial rough presentation content

@@
 
+# Research Question
+
+- SolarStat: Modeling Photovolatic Sources through Stochastic Markov Processes
+    - Marco Miozzo, Davide Zordan, Paolo Dini, and Michele Rossi
+- Their objective: Describe how solar irradiance is translated into current output in a PV system throughout the day and cluster results with a Markov model to determine day/night breaks.
+- Our objective: Describe how solar irradiance is translated into current output in a PV system and bootsrap results both with and without respect to time, and compare.
+
 # Background
+
+- Understanding renewable energy sources like solar power is critical to the next century. 
+- PV modules installed at individual households/nodes reduce the overall load on the electrical grid. 
+- Balancing the expected contribution of an individual PV module with the expected residual demand required of the grid is a large optimization problem. 
+    - We dealt with half of that optimization problem involving the expected contribution of the PV module.
+
+# Data
+
+- National Solar Radiation Database (NSRD) from National Renewable Energy Laboratory (NREL) accessed through API
+- Los Angeles data (1998-2010) was used to replicate the source paper. (IS THE YEAR RANGE RIGHT BOSTON?)
+    - Date and time, direct normal irradiance (DNI), and temperature data used.
+
+# Preprocessing
+
+INSERT FIG 1 FROM PAPER HERE
+- Astronomical adjustments performed on DNI to find effective solar irradiance:
+    - Latitude, longitude, time zone, and the tilt of the earth used in calculations. 
+    - $I_{eff} \propto I_{sun}cos(\theta)$
+
+# Photovolatic (PV) Module
+
+- Power processing units are attached to PV modules to optimize power output for a given current input.
+    - Maximizes power $P=IV$ over a range of possible voltages. 
+- Result of interest is current output after being adjusted by the power processor and an assumed leakage rate.
+
+# Bootstrapping
+
+# Results
+
+# Future Research
+
+- A similar procedure could be performed on a different city's data like Fort Collins.
+- A similar procedure could be performed over a different range of years. A comparison may reveal the effects of global warming. 
+- Our original source paper examined multiple PV modules simultaneously in the context of a larger grid. This examines both sides of the contribution/demand optimization problem.

### 2021-12-06 deliverables/presentation.Rmd
Slight rearrangement, dropped trailing whitespace

@@
   beamer_presentation: default
+bibliography: references.bib
 ---
@@
 
-# Background
+# Motivation
 
@@
 
+# Background
+
 # Data
@@
 - National Solar Radiation Database (NSRD) from National Renewable Energy Laboratory (NREL) accessed through API
-- Los Angeles data (1998-2010) was used to replicate the source paper. (IS THE YEAR RANGE RIGHT BOSTON?)
+- Los Angeles data (1998-2010) was used to replicate the source paper.
     - Date and time, direct normal irradiance (DNI), and temperature data used.

### 2021-12-06 deliverables/paper.Rmd
Added EDA functions, day segmenting code

--- 2021-12-01 deliverables/paper.Rmd
@@
 ```
+
+```{r data}
+df <- read_csv("../data/final/irradiance_full_final.csv")
+```
+
+# EDA
+
+```{r timeseries}
+
+#' Plot a timeseries of a single variable
+#'
+#' @param df A dataframe containing `variable`
+#' @param variable A string referring to a column of `df`
+#' @param n_days The number of days (starting from the
+#'    beginning of the data) to plot
+#' @return A ggplot object of a timeseries of `variable` for `n_days` days
+time_series_of_var <- function(df, variable, n_days) {
+ret_plot <- df %>%
+  mutate(date = lubridate::make_datetime(Year, Month, Day, Hour, Minute)) %>%
+  head(24 * n_days) %>%
+  ggplot() +
+  geom_line(aes_string(x = "date", y = variable)) +
+  theme_bw()
+return(ret_plot)
+}
+
+time_series_of_var(df, "GHI", 6)
+time_series_of_var(df, "current_out", 6)
+```
+
+```{r hour-plot}
+
+#' Plot the values of `variable` at each hour of the day
+#'
+#' @param df A dataframe containing `variable`
+#' @param variable A string referring to a column of `df`
+#' @return A ggplot object of a boxplot of `variable` segmented by hour
+hour_plot_of_var <- function(df, variable) {
+df$Hour <- as.factor(df$Hour)
+ret_plot <- df %>%
+  ggplot() +
+  geom_boxplot(aes_string(x = "Hour", y = variable)) +
+  xlab("Hour (0-23)") +
+  ylab(paste("Distribution of", variable)) +
+  theme_bw()
+return(ret_plot)
+}
+
+hour_plot_of_var(df, "GHI")
+hour_plot_of_var(df, "current_out")
+```
+
+# Data filtering
+
+```{r filter}
+HOUR_WINDOW <- 11:16
+df_high_irradiance <- df[df$Hour %in% HOUR_WINDOW,]
+filtered_current_hist <- ggplot() +
+  geom_histogram(aes(x = df_high_irradiance$current_out)) +
+  theme_bw()
+print(filtered_current_hist)
+```
+
+# Bootstrapping
+
+```{r functions}
+
+np <- function(x, l) {
+  n <- length(x)
+  print(n)
+  # l <- n %/% b
+  # b <- n %/% l
+  b <- floor(n / l)
+  print(b)
+  print(n / b)
+  blocks <- matrix(NA, nrow = b, ncol = l)
+  for (i in 1:b) {
+    # blocks[i, ] <- x[(i - 1) * l + 1:(i * l)]
+    print(x[(i - 1) * l + 1:(i * l)])
+  }
+  return(blocks)
+}
+
+nbb <- function(x, l) {
+  B <- 10000
+  b <- floor(length(x) / l)
+  theta_hat_star_nbb <- rep(NA, B)
+  # nb_blocks <- np(x, l, b)
+  nb_blocks <- np(x, l)
+  for (i in 1:B) {
+    # sample blocks
+    idx <- sample(1:b, b, replace = TRUE)
+    theta_hat_star_nbb[i] <- mean(x[idx, ])
+  }
+  ggplot() +
+    geom_histogram(aes(theta_hat_star_nbb))
+
+  st.d <- sd(theta_hat_star_nbb)
+  bias <- mean(theta_hat_star_nbb) - mean(x)
+
+  print(paste(
+    "The standard deviation with ", l, " blocks is: ", st.d, ". \n The bias is: ",
+    bias
+  ))
+}
+```
+
+```{r sampling}
+# B1 <- df[df$Hour == c(7:10),]$current_output # Hours before peak
+# B2 <- df[df$Hour == c(11:14),]$current_output # Peak hours
+# B3 <- df[df$Hour == c(15:18),]$current_output # Hours after peak
+
+nbb(hours, 4) # in order to make b blocks I would need to solve for b = [n/l] (floor) where n is the length of our data and l is 4, so make nbb call l instead of b
+
+nbb(hours, 12)
+
+# Could also block instead by hours but by total current output.
+Peak hours would be a frame around the highest output and then hours before and after would coincide with that.
+This would cause the hours to potentially be different based on the season/month as months that get dark earlier could see peak hours earlier in the day.
+
+# dpeak <- c(B1,B2,B3)
+```

### 2021-12-06 deliverables/paper.Rmd
Added function for creating bootstrap blocks from df

@@
 HOUR_WINDOW <- 11:16
-df_high_irradiance <- df[df$Hour %in% HOUR_WINDOW,]
+df_high_irradiance <- df[df$Hour %in% HOUR_WINDOW, ]
 filtered_current_hist <- ggplot() +
@@
 ```{r functions}
+
+#' Given a dataframe, return a matrix of bootstrap blocks
+#'
+#' @param df A dataframe of irradiance values
+#' @return A matrix whose rows are bootstrap blocks
+make_blocks_from_df <- function(df) {
+  df <- df %>%
+    mutate(date = lubridate::make_date(Year, Month, Day))
+  # The number of blocks is the number of days
+  number_of_blocks <- length(unique(df$date))
+  # The length of a block is the number of
+  # observations in each individual day.
+  # In this case, we can simply look at the first one,
+  # because all of the days have the same number of observations
+  length_of_block <- nrow(df[df$date == unique(df$date)[1], ])
+  blocks <- matrix(NA, nrow = number_of_blocks, ncol = length_of_block)
+  print(dim(blocks))
+  for (i in 1:number_of_blocks) {
+    blocks[i, ] <- df$current_out[(i - 1) * length_of_block + 1 : (i * length_of_block)]
+  }
+  return(blocks)
+}
+# We have to use the modified DataFrame here,
+# or else the assumptions about block size don't hold
+make_blocks_from_df(df_high_irradiance)
 

### 2021-12-06 data_preprocessing/main.py
Changed to correct radiance measure

@@ -20,7 +20,7 @@ def main():
 
     df[SOLAR_IRRADIANCE] = df.apply(
         lambda row: aa.effective_solar_radiance(
-            row["DHI"],
+            row["GHI"],
             row["Month"],
             row["Day"],
             row["Hour"],

### 2021-12-07 data_preprocessing/astronomical_adjustment.py
Fixed minor errors in astronomical calculation

@@ -53,7 +53,7 @@ class AstronomicalAdjustment:
         """
         EARTH_ANGLE = math.radians(23.45)
         D = cls.d_helper(day_number)
-        return np.arcsin(math.sin(EARTH_ANGLE) * math.sin(D))
+        return math.asin(math.sin(EARTH_ANGLE) * math.sin(D))
 
     @classmethod
     def hour_angle(cls, time, day_number, longitude, utc_off):
@@ -70,7 +70,8 @@ class AstronomicalAdjustment:
 
         """
         hour_angle_degrees = (
-            15 * cls.apparent_solar_time(time, day_number, longitude, utc_off) - 12
+            15 * cls.apparent_solar_time(time, day_number, longitude, utc_off)
+            - 12
         )
         return math.radians(hour_angle_degrees)
 
@@ -112,7 +113,9 @@ class AstronomicalAdjustment:
 
         """
         D = cls.d_helper(day_number)
-        return (9.87 * math.sin(2 * D) - 7.53 * math.cos(D) - 1.5 * math.sin(D)) / 60
+        return (
+            9.87 * math.cos(2 * D) - 7.53 * math.cos(D) - 1.5 * math.sin(D)
+        ) / 60
 
     @classmethod
     def angle_sun_surface(cls, time, day_number, latitude, longitude, utc_off):
@@ -136,7 +139,7 @@ class AstronomicalAdjustment:
         BETA = cls.BETA
         declination_angle = cls.declination_angle(day_number)
         hour_angle = cls.hour_angle(time, day_number, longitude, utc_off)
-        latitude = np.radians(latitude)
+        latitude = abs(math.radians(latitude))
         ret_val = (
             (math.sin(declination_angle) * math.sin(latitude) * math.cos(BETA))
             - (
@@ -199,7 +202,7 @@ class AstronomicalAdjustment:
         """
 
         cumulative_days = [0, 31, 28, 31, 30, 31, 30, 31, 31, 30, 31, 30, 31]
-        return sum(cumulative_days[0 : int(month - 1)]) + day
+        return sum(cumulative_days[0 : int(month)]) + day
 
     @classmethod
     def effective_solar_radiance(
@@ -229,5 +232,7 @@ class AstronomicalAdjustment:
 
         return i_sun * max(
             0,
-            cls.angle_sun_surface(time, day_number, latitude, longitude, utc_off),
+            cls.angle_sun_surface(
+                time, day_number, latitude, longitude, utc_off
+            ),
         )

### 2021-12-07 deliverables/paper.Rmd
Cleaned up bootstrapping code, not quite operational

@@
 hour_plot_of_var(df, "current_out")
+hour_plot_of_var(df, "declination_angle")
+hour_plot_of_var(df, "solar_time")
+hour_plot_of_var(df, "angle_sun_surface_degrees")
 ```
@@
   blocks <- matrix(NA, nrow = number_of_blocks, ncol = length_of_block)
-  print(dim(blocks))
+  start <- 1
+  stop <- length_of_block
   for (i in 1:number_of_blocks) {
-    blocks[i, ] <- df$current_out[(i - 1) * length_of_block + 1 : (i * length_of_block)]
+    # blocks[i, ] <- df$current_out[(i - 1) * length_of_block + 1 : (i * length_of_block)]
+    blocks[i, ] <- df$current_out[start:stop]
+    start <- start + length_of_block
+    stop <- stop + length_of_block
   }
@@
 # or else the assumptions about block size don't hold
-make_blocks_from_df(df_high_irradiance)
+test <- make_blocks_from_df(df_high_irradiance)
+
+block_bootstrap_estimates_from_df <- function(df, B=10000) {
+  block_bootstrap_estimates <- rep(NA, B)
+  blocks <- make_blocks_from_df(df)
+  number_of_blocks <- nrow(blocks)
+  for (i in 1:B) {
+    # sample blocks
+    idx <- sample(1:number_of_blocks, number_of_blocks, replace = TRUE)
+    # Concatenate blocks at the given indices
+    # This gives a "pseudo-dataset" containing resampled blocks
+    block_bootstrap_estimates[i] <- mean(as.vector(t(blocks[idx, ])))
+  }
+  return(block_bootstrap_estimates)
+}
+test_est <- block_bootstrap_estimates_from_df(df_high_irradiance)
 

### 2021-12-07 deliverables/paper.Rmd
Added initial comparison to normal bootstrap

@@
 hour_plot_of_var(df, "current_out")
-hour_plot_of_var(df, "declination_angle")
-hour_plot_of_var(df, "solar_time")
-hour_plot_of_var(df, "angle_sun_surface_degrees")
 ```
@@
     # This gives a "pseudo-dataset" containing resampled blocks
-    block_bootstrap_estimates[i] <- mean(as.vector(t(blocks[idx, ])))
+    block_bootstrap_estimates[i] <- median(as.vector(t(blocks[idx, ])))
   }
@@
 test_est <- block_bootstrap_estimates_from_df(df_high_irradiance)
+
+meanv <- function(data, idx) {
+  return(mean(data[idx]))
+}
+
+sdv <- function(data, idx) {
+  return(sd(data[idx]))
+}
+
+medianv <- function(data, idx) {
+  return(median(data[idx]))
+}
+
+iid_bootstrap_estimates_from_df <- function(df, B=10000) {
+  boot_current <- boot(df$current_out, medianv, B)
+  return(boot_current$t)
+}
+
+test_iid <- iid_bootstrap_estimates_from_df(df_high_irradiance)
+
+ggplot() +
+  geom_histogram(aes(x = test_est),color="red",alpha=0.5) +
+  geom_histogram(aes(x = test_iid),color="blue",alpha=0.5) +
+  theme_bw()
 

### 2021-12-07 data_preprocessing/astronomical_adjustment.py
Fixed minor mistakes in astronomical adjustment

@@ -69,9 +69,8 @@ class AstronomicalAdjustment:
             float: Returns the hour angle in radians
 
         """
-        hour_angle_degrees = (
-            15 * cls.apparent_solar_time(time, day_number, longitude, utc_off)
-            - 12
+        hour_angle_degrees = 15 * (
+            cls.apparent_solar_time(time, day_number, longitude, utc_off) - 12
         )
         return math.radians(hour_angle_degrees)
 
@@ -114,7 +113,7 @@ class AstronomicalAdjustment:
         """
         D = cls.d_helper(day_number)
         return (
-            9.87 * math.cos(2 * D) - 7.53 * math.cos(D) - 1.5 * math.sin(D)
+            9.87 * math.sin(2 * D) - 7.53 * math.cos(D) - 1.5 * math.sin(D)
         ) / 60
 
     @classmethod

### 2021-12-08 deliverables/paper.Rmd
Added images subdirectory and preliminary figure saving

@@
 library(tidyverse)
+library(boot)
 ```
@@
 
-block_bootstrap_estimates_from_df <- function(df, B=10000) {
+block_bootstrap_estimates_from_df <- function(df, B = 10000) {
   block_bootstrap_estimates <- rep(NA, B)
@@
     # This gives a "pseudo-dataset" containing resampled blocks
-    block_bootstrap_estimates[i] <- median(as.vector(t(blocks[idx, ])))
+    block_bootstrap_estimates[i] <- mean(as.vector(t(blocks[idx, ])))
   }
@@
 
-iid_bootstrap_estimates_from_df <- function(df, B=10000) {
-  boot_current <- boot(df$current_out, medianv, B)
+iid_bootstrap_estimates_from_df <- function(df, B = 10000) {
+  boot_current <- boot(df$current_out, meanv, B)
   return(boot_current$t)
@@
 
-ggplot() +
-  geom_histogram(aes(x = test_est),color="red",alpha=0.5) +
-  geom_histogram(aes(x = test_iid),color="blue",alpha=0.5) +
+mean_samp_dist <- ggplot() +
+  geom_histogram(aes(x = test_est), color = "red", alpha = 0.5) +
+  geom_histogram(aes(x = test_iid), color = "blue", alpha = 0.5) +
   theme_bw()
-
-np <- function(x, l) {
-  n <- length(x)
-  print(n)
-  # l <- n %/% b
-  # b <- n %/% l
-  b <- floor(n / l)
-  print(b)
-  print(n / b)
-  blocks <- matrix(NA, nrow = b, ncol = l)
-  for (i in 1:b) {
-    # blocks[i, ] <- x[(i - 1) * l + 1:(i * l)]
-    print(x[(i - 1) * l + 1:(i * l)])
-  }
-  return(blocks)
-}
-
-nbb <- function(x, l) {
-  B <- 10000
-  b <- floor(length(x) / l)
-  theta_hat_star_nbb <- rep(NA, B)
-  # nb_blocks <- np(x, l, b)
-  nb_blocks <- np(x, l)
-  for (i in 1:B) {
-    # sample blocks
-    idx <- sample(1:b, b, replace = TRUE)
-    theta_hat_star_nbb[i] <- mean(x[idx, ])
-  }
-  ggplot() +
-    geom_histogram(aes(theta_hat_star_nbb))
-
-  st.d <- sd(theta_hat_star_nbb)
-  bias <- mean(theta_hat_star_nbb) - mean(x)
-
-  print(paste(
-    "The standard deviation with ", l, " blocks is: ", st.d, ". \n The bias is: ",
-    bias
-  ))
-}
+ggsave("./images/mean_samp_dist.png", mean_samp_dist)
 ```
@@
 
-nbb(hours, 4) # in order to make b blocks I would need to solve for b = [n/l] (floor) where n is the length of our data and l is 4, so make nbb call l instead of b
+# nbb(hours, 4) # in order to make b blocks I would need to solve for b = [n/l] (floor) where n is the length of our data and l is 4, so make nbb call l instead of ib
 
-nbb(hours, 12)
+# nbb(hours, 12)
 
-# Could also block instead by hours but by total current output.
-Peak hours would be a frame around the highest output and then hours before and after would coincide with that.
-This would cause the hours to potentially be different based on the season/month as months that get dark earlier could see peak hours earlier in the day.
+# Could also block instead by hours but by total current output. Peak hours
+# would be a frame around the highest output and then hours before and after
+# would coincide with that. This would cause the hours to potentially be
+# different based on the season/month as months that get dark earlier could see
+# peak hours earlier in the day.
 

### 2021-12-08 deliverables/presentation.Rmd
Added notes on bootstrapping

@@
 
+- We chose our initial paper because of its use of Monte Carlo.
+After working with it for a little it was brought to our attention that Monte Carlo was used in conjunction with a backward/forward sweep method in order to calculate the pdf.
+After some confusion it was decided to switch to another paper dealing with the same data/motivation but without said methods (including Monte Carlo).
+- Luckily, the week after we switched papers we begun learning about bootstrapping, a way to sample from our data while taking into consideration our time change.
+
+INSERT GRAPHS FROM IID BOOTSTRAPPING AND BLOCK HERE
+ALSO INSERT FIG 1 FROM PAPER
+
+- The paper focused most of their sampling on night-day clustering, showing the difference between night and day through current.
+Night yielded thousands of 0 current values, leading to skewed data when sampling.
+Instead we took too approaches to bootstrap the current vs.
+hour data.
+- One using non-overlapping blocks sampling from the 8 hours surrounding the peak irradiance. The other being IID, yielding dependence in our data.
+
 # Results

### 2021-12-08 deliverables/presentation.Rmd
Simplified presentation slides

--- 2021-12-06 deliverables/presentation.Rmd
@@
 
-- SolarStat: Modeling Photovolatic Sources through Stochastic Markov Processes
-    - Marco Miozzo, Davide Zordan, Paolo Dini, and Michele Rossi
-- Their objective: Describe how solar irradiance is translated into current output in a PV system throughout the day and cluster results with a Markov model to determine day/night breaks.
-- Our objective: Describe how solar irradiance is translated into current output in a PV system and bootsrap results both with and without respect to time, and compare.
+- _SolarStat: Modeling Photovolatic Sources through Stochastic Markov Processes_
+
+- *Their objective*: 
+  - Solar irradiance $\rightarrow$ current output 
+  - Cluster results with a Markov model
+- *Our objective*: 
+  - Solar irradiance $\rightarrow$ current output 
+  - Compare iid and time series bootstrap
 
@@
 - Balancing the expected contribution of an individual PV module with the expected residual demand required of the grid is a large optimization problem.
-    - We dealt with half of that optimization problem involving the expected contribution of the PV module.
+    - We dealt with half of that optimization problem
 
@@
 
+- The paper used data from LA, 1991-2010
+- The paper uses ETR (we used similar GHI)
+- Simulates for a specific model of small solar panel 
+- Panels "laid out" in a 6x6 grid.
+
 # Data
 
-- National Solar Radiation Database (NSRD) from National Renewable Energy Laboratory (NREL) accessed through API
-- Los Angeles data (1998-2010) was used to replicate the source paper.
-    - Date and time, direct normal irradiance (DNI), and temperature data used.
+- National Solar Radiation Database (NSRDB) from National Renewable Energy
+Laboratory (NREL)
+- Los Angeles data for 1998-2010
 
-# Preprocessing
+# Overall process
+
+\begin{center}
+Raw data (NSRDB)
+
+$\downarrow$
+
+Astronomical adjustment
+
+$\downarrow$
+
+PV module
+
+$\downarrow$
+
+Power processor
+
+$\downarrow$
+
+Bootstrap
+\end{center}
+
+# Preprocessing: Astronomical Adjustment
 
 INSERT FIG 1 FROM PAPER HERE
-- Astronomical adjustments performed on DNI to find effective solar irradiance:
-    - Latitude, longitude, time zone, and the tilt of the earth used in calculations.
-    - $I_{eff} \propto I_{sun}cos(\theta)$
+- Astronomical adjustments performed on DNI to find *effective* solar irradiance
+- Used: Latitude, longitude, time zone, tilt of the earth 
+- $I_{eff} \propto I_{sun}cos(\Theta)$
 
-# Photovolatic (PV) Module
+# Preprocessing: Power processor
+
+- This gives a procedure for translating effective irradiance to
+current, given an operating voltage
+- Uses:
+  - Effective irradiance
+  - Operating voltage
+  - Properties of solar panel materials
+  - Solar panel layout
+
+# Preprocessing: Photovoltaic (PV) Module
 
 - Power processing units are attached to PV modules to optimize power output for a given current input.
-    - Maximizes power $P=IV$ over a range of possible voltages.
-- Result of interest is current output after being adjusted by the power processor and an assumed leakage rate.
+- Maximizes power $P=IV$ over voltage
+- Here $I$ is a function of $V$
+- Result: Optimal current
+
+# Results
+
+ADD FIGURES
 
 # Bootstrapping
-
-# Results
 

### 2021-12-08 deliverables/paper.Rmd
Added Evan's notes to paper draft

@@
 library(tidyverse)
-library(boot)
 ```
 
-```{r data}
-df <- read_csv("../data/final/irradiance_full_final.csv")
-```
+- We chose our initial paper because of its use of Monte Carlo.
+After working with it for a little it was brought to our attention that Monte Carlo was used in conjunction with a backward/forward sweep method in order to calculate the pdf.
+After some confusion it was decided to switch to another paper dealing with the same data/motivation but without said methods (including Monte Carlo).
+- Luckily, the week after we switched papers we begun learning about bootstrapping, a way to sample from our data while taking into consideration our time change.
 
-# EDA
+INSERT GRAPHS FROM IID BOOTSTRAPPING AND BLOCK HERE
+ALSO INSERT FIG 1 FROM PAPER
 
-```{r timeseries}
-
-#' Plot a timeseries of a single variable
-#'
-#' @param df A dataframe containing `variable`
-#' @param variable A string referring to a column of `df`
-#' @param n_days The number of days (starting from the
-#'    beginning of the data) to plot
-#' @return A ggplot object of a timeseries of `variable` for `n_days` days
-time_series_of_var <- function(df, variable, n_days) {
-  ret_plot <- df %>%
-    mutate(date = lubridate::make_datetime(Year, Month, Day, Hour, Minute)) %>%
-    head(24 * n_days) %>%
-    ggplot() +
-    geom_line(aes_string(x = "date", y = variable)) +
-    theme_bw()
-  return(ret_plot)
-}
-
-time_series_of_var(df, "GHI", 6)
-time_series_of_var(df, "current_out", 6)
-```
-
-```{r hour-plot}
-
-#' Plot the values of `variable` at each hour of the day
-#'
-#' @param df A dataframe containing `variable`
-#' @param variable A string referring to a column of `df`
-#' @return A ggplot object of a boxplot of `variable` segmented by hour
-hour_plot_of_var <- function(df, variable) {
-  df$Hour <- as.factor(df$Hour)
-  ret_plot <- df %>%
-    ggplot() +
-    geom_boxplot(aes_string(x = "Hour", y = variable)) +
-    xlab("Hour (0-23)") +
-    ylab(paste("Distribution of", variable)) +
-    theme_bw()
-  return(ret_plot)
-}
-
-hour_plot_of_var(df, "GHI")
-hour_plot_of_var(df, "current_out")
-```
-
-# Data filtering
-
-```{r filter}
-HOUR_WINDOW <- 11:16
-df_high_irradiance <- df[df$Hour %in% HOUR_WINDOW, ]
-filtered_current_hist <- ggplot() +
-  geom_histogram(aes(x = df_high_irradiance$current_out)) +
-  theme_bw()
-print(filtered_current_hist)
-```
-
-# Bootstrapping
-
-```{r functions}
-
-#' Given a dataframe, return a matrix of bootstrap blocks
-#'
-#' @param df A dataframe of irradiance values
-#' @return A matrix whose rows are bootstrap blocks
-make_blocks_from_df <- function(df) {
-  df <- df %>%
-    mutate(date = lubridate::make_date(Year, Month, Day))
-  # The number of blocks is the number of days
-  number_of_blocks <- length(unique(df$date))
-  # The length of a block is the number of
-  # observations in each individual day.
-  # In this case, we can simply look at the first one,
-  # because all of the days have the same number of observations
-  length_of_block <- nrow(df[df$date == unique(df$date)[1], ])
-  blocks <- matrix(NA, nrow = number_of_blocks, ncol = length_of_block)
-  start <- 1
-  stop <- length_of_block
-  for (i in 1:number_of_blocks) {
-    # blocks[i, ] <- df$current_out[(i - 1) * length_of_block + 1 : (i * length_of_block)]
-    blocks[i, ] <- df$current_out[start:stop]
-    start <- start + length_of_block
-    stop <- stop + length_of_block
-  }
-  return(blocks)
-}
-# We have to use the modified DataFrame here,
-# or else the assumptions about block size don't hold
-test <- make_blocks_from_df(df_high_irradiance)
-
-block_bootstrap_estimates_from_df <- function(df, B = 10000) {
-  block_bootstrap_estimates <- rep(NA, B)
-  blocks <- make_blocks_from_df(df)
-  number_of_blocks <- nrow(blocks)
-  for (i in 1:B) {
-    # sample blocks
-    idx <- sample(1:number_of_blocks, number_of_blocks, replace = TRUE)
-    # Concatenate blocks at the given indices
-    # This gives a "pseudo-dataset" containing resampled blocks
-    block_bootstrap_estimates[i] <- mean(as.vector(t(blocks[idx, ])))
-  }
-  return(block_bootstrap_estimates)
-}
-test_est <- block_bootstrap_estimates_from_df(df_high_irradiance)
-
-meanv <- function(data, idx) {
-  return(mean(data[idx]))
-}
-
-sdv <- function(data, idx) {
-  return(sd(data[idx]))
-}
-
-medianv <- function(data, idx) {
-  return(median(data[idx]))
-}
-
-iid_bootstrap_estimates_from_df <- function(df, B = 10000) {
-  boot_current <- boot(df$current_out, meanv, B)
-  return(boot_current$t)
-}
-
-test_iid <- iid_bootstrap_estimates_from_df(df_high_irradiance)
-
-mean_samp_dist <- ggplot() +
-  geom_histogram(aes(x = test_est), color = "red", alpha = 0.5) +
-  geom_histogram(aes(x = test_iid), color = "blue", alpha = 0.5) +
-  theme_bw()
-ggsave("./images/mean_samp_dist.png", mean_samp_dist)
-```
-
-```{r sampling}
-# B1 <- df[df$Hour == c(7:10),]$current_output # Hours before peak
-# B2 <- df[df$Hour == c(11:14),]$current_output # Peak hours
-# B3 <- df[df$Hour == c(15:18),]$current_output # Hours after peak
-
-# nbb(hours, 4) # in order to make b blocks I would need to solve for b = [n/l] (floor) where n is the length of our data and l is 4, so make nbb call l instead of ib
-
-# nbb(hours, 12)
-
-# Could also block instead by hours but by total current output. Peak hours
-# would be a frame around the highest output and then hours before and after
-# would coincide with that. This would cause the hours to potentially be
-# different based on the season/month as months that get dark earlier could see
-# peak hours earlier in the day.
-
-# dpeak <- c(B1,B2,B3)
-```
+- The paper focused most of their sampling on night-day clustering, showing the difference between night and day through current.
+Night yielded thousands of 0 current values, leading to skewed data when sampling.
+Instead we took too approaches to bootstrap the current vs.
+hour data.
+- One using non-overlapping blocks sampling from the 8 hours surrounding the peak irradiance. The other being IID, yielding dependence in our data.

### 2021-12-08 deliverables/presentation.Rmd
Simplified preprocessing slides

@@
 
-# Preprocessing: Astronomical Adjustment
+# Preprocessing
 
-INSERT FIG 1 FROM PAPER HERE
-- Astronomical adjustments performed on DNI to find *effective* solar irradiance
-- Used: Latitude, longitude, time zone, tilt of the earth 
-- $I_{eff} \propto I_{sun}cos(\Theta)$
+1. Astronomical adjustment: GHI $\rightarrow$ effective irradiance
+1. Power processor: Effective irradiance + voltage $\rightarrow$ current 
+1. PV Module: Effective irradiance $\rightarrow$ current
 
-# Preprocessing: Power processor
-
-- This gives a procedure for translating effective irradiance to
-current, given an operating voltage
-- Uses:
-  - Effective irradiance
-  - Operating voltage
-  - Properties of solar panel materials
-  - Solar panel layout
-
-# Preprocessing: Photovoltaic (PV) Module
-
-- Power processing units are attached to PV modules to optimize power output for a given current input.
-- Maximizes power $P=IV$ over voltage
-- Here $I$ is a function of $V$
-- Result: Optimal current
-
-# Results
+# Postprocessed data
 

### 2021-12-08 deliverables/paper.Rmd

--- 2021-12-08 deliverables/paper.Rmd
@@
 ```
+
+- We chose our initial paper because of its use of Monte Carlo.
+After working with it for a little it was brought to our attention that Monte Carlo was used in conjunction with a backward/forward sweep method in order to calculate the pdf.
+After some confusion it was decided to switch to another paper dealing with the same data/motivation but without said methods (including Monte Carlo).
+- Luckily, the week after we switched papers we begun learning about bootstrapping, a way to sample from our data while taking into consideration our time change.
+
+INSERT GRAPHS FROM IID BOOTSTRAPPING AND BLOCK HERE
+ALSO INSERT FIG 1 FROM PAPER
+
+- The paper focused most of their sampling on night-day clustering, showing the difference between night and day through current.
+Night yielded thousands of 0 current values, leading to skewed data when sampling.
+Instead we took too approaches to bootstrap the current vs.
+hour data.
+- One using non-overlapping blocks sampling from the 8 hours surrounding the peak irradiance. The other being IID, yielding dependence in our data.

### 2021-12-08 deliverables/paper.Rmd
Added figure-saving to paper and figures to presentation

@@
 
+```{r plot-pres-func}
+make_plot_for_pres <- function(plot) {
+  ret_plot <- plot +
+  theme(text=element_text(size=21))
+  return(ret_plot)
+}
+```
+
 # EDA
@@
 #' @param variable A string referring to a column of `df`
+#' @param var_name The human-readable name of `variable`, for labelling
 #' @param n_days The number of days (starting from the
@@
 #' @return A ggplot object of a timeseries of `variable` for `n_days` days
-time_series_of_var <- function(df, variable, n_days) {
+time_series_of_var <- function(df, variable, var_name, n_days) {
   ret_plot <- df %>%
@@
     geom_line(aes_string(x = "date", y = variable)) +
+    xlab("Date") +
+    ylab(var_name) +
     theme_bw()
@@
 
-time_series_of_var(df, "GHI", 6)
-time_series_of_var(df, "current_out", 6)
+time_series_ghi <- time_series_of_var(df, "GHI", "GHI", 6)
+time_series_current <- time_series_of_var(df, "current_out", "Output current", 6)
+ggsave("./images/time_series_ghi.png", make_plot_for_pres(time_series_ghi))
+ggsave("./images/time_series_current.png", make_plot_for_pres(time_series_current))
 ```
@@
 #' @param variable A string referring to a column of `df`
+#' @param var_name The human-readable name of `variable`, for labelling
 #' @return A ggplot object of a boxplot of `variable` segmented by hour
-hour_plot_of_var <- function(df, variable) {
+hour_plot_of_var <- function(df, variable, var_name) {
   df$Hour <- as.factor(df$Hour)
@@
     xlab("Hour (0-23)") +
-    ylab(paste("Distribution of", variable)) +
+    ylab(paste("Distribution of", var_name)) +
     theme_bw()
@@
 
-hour_plot_of_var(df, "GHI")
-hour_plot_of_var(df, "current_out")
+hour_plot_ghi <- hour_plot_of_var(df, "GHI", "GHI")
+hour_plot_current <- hour_plot_of_var(df, "current_out", "output current")
+ggsave("./images/hour_plot_ghi.png", make_plot_for_pres(hour_plot_ghi))
+ggsave("./images/hour_plot_current.png", make_plot_for_pres(hour_plot_current))
 ```

### 2021-12-08 deliverables/presentation.Rmd
Fixed preprocessing order
Added figure-saving to paper and figures to presentation

@@
 
+::: columns
+:::: column
+![](./images/time_series_ghi.png)
+::::
+:::: column
+![](./images/hour_plot_ghi.png)
+::::
+:::
+
 - National Solar Radiation Database (NSRDB) from National Renewable Energy
@@
 1. Astronomical adjustment: GHI $\rightarrow$ effective irradiance
-1. Power processor: Effective irradiance + voltage $\rightarrow$ current 
-1. PV Module: Effective irradiance $\rightarrow$ current
+1. PV Module: Effective irradiance + voltage $\rightarrow$ current 
+1. Power processor: Effective irradiance $\rightarrow$ current
 
@@
 
-ADD FIGURES
+::: columns
+:::: column
+![](./images/time_series_ghi.png)
+::::
+:::: column
+![](./images/time_series_current.png)
+::::
+:::
+
+# Postprocessed data, cont.
+
+::: columns
+:::: column
+![](./images/hour_plot_ghi.png)
+::::
+:::: column
+![](./images/hour_plot_current.png)
+::::
+:::
 

### 2021-12-08 deliverables/paper.Rmd
Added tick labels to fit image resizing

@@
     ylab(paste("Distribution of", var_name)) +
+    scale_x_discrete(breaks=seq(0,23,2)) +
     theme_bw()

### 2021-12-08 deliverables/paper.Rmd
Added code for generating bootstrap histogram, and histogram in presentation

--- 2021-12-08 deliverables/paper.Rmd
@@
   ret_plot <- plot +
-  theme(text=element_text(size=21))
+    theme(text = element_text(size = 21))
   return(ret_plot)
@@
     ylab(paste("Distribution of", var_name)) +
+    scale_x_discrete(breaks = seq(0, 23, 2)) +
     theme_bw()
@@
 }
-test_est <- block_bootstrap_estimates_from_df(df_high_irradiance)
 
@@
 
-test_iid <- iid_bootstrap_estimates_from_df(df_high_irradiance)
+boot_block <- block_bootstrap_estimates_from_df(df_high_irradiance)
+boot_block_df <- data.frame(boot = boot_block, type = rep("Block", length(boot_block)))
+boot_iid <- iid_bootstrap_estimates_from_df(df_high_irradiance)
+boot_iid_df <- data.frame(boot = boot_iid, type = rep("iid", length(boot_iid)))
 
-mean_samp_dist <- ggplot() +
-  geom_histogram(aes(x = test_est), color = "red", alpha = 0.5) +
-  geom_histogram(aes(x = test_iid), color = "blue", alpha = 0.5) +
+vis_df <- rbind(boot_block_df, boot_iid_df)
+
+mean_boot_plot <- ggplot(vis_df) +
+  geom_histogram(aes(x = boot, y=..density.., fill = type), position = "identity", alpha = 0.5, color = "black") +
+  xlab("Bootstrap sample means") +
+  labs(fill = "Bootstrap type") +
   theme_bw()
-ggsave("./images/mean_samp_dist.png", mean_samp_dist)
+ggsave("./images/mean_boot_plot.png", make_plot_for_pres(mean_boot_plot), width = 16, height = 12)
 ```

### 2021-12-08 deliverables/paper.Rmd
Bootstrap CIs and MC coverage estimate

@@
 
-```{r libs}
+```{r libs, include=FALSE, message=FALSE}
 library(tidyverse)
@@
 
-```{r data}
-df <- read_csv("../data/final/irradiance_full_final.csv")
+```{r data, message=FALSE, warning=FALSE}
+df <- read_csv("../data/final/irradiance_full_final.csv") %>%
+  select(-'...1')
 ```
@@
 
-```{r timeseries}
+```{r timeseries, warning=FALSE}
 
@@
 filtered_current_hist <- ggplot() +
-  geom_histogram(aes(x = df_high_irradiance$current_out)) +
-  theme_bw()
+  geom_histogram(aes(x = df_high_irradiance$current_out, y = ..density..), binwidth = .0025) +
+  theme_bw() +
+  labs(title='Distribution of Current Outputs', subtitle='Restricted to high irradiance hours of the day', 
+       x = 'Current Output', y = 'Density')
 print(filtered_current_hist)
@@
 
+# percentile bootstrap CI used
+alpha <- .05
+block_ci <- quantile(boot_block_df$boot, c(alpha / 2, 1 - alpha / 2))
+iid_ci <- quantile(boot_iid_df$boot, c(alpha / 2, 1 - alpha / 2))
+
 mean_boot_plot <- ggplot(vis_df) +
-  geom_histogram(aes(x = boot, y=..density.., fill = type), position = "identity", alpha = 0.5, color = "black") +
-  xlab("Bootstrap sample means") +
-  labs(fill = "Bootstrap type") +
+  geom_histogram(aes(x = boot, y = ..density.., fill = type), position = "identity", alpha = .5, color = "black", 
+                 binwidth = .0001) +
+  # geom_segment(aes(x = block_ci[1], xend = block_ci[2], y = 1250, yend = 1250), color = 'tomato', alpha = .01) +
+  geom_vline(aes(xintercept = block_ci[1]), linetype = 'dashed', color = 'tomato', alpha = .5) +
+  geom_vline(aes(xintercept = block_ci[2]), linetype = 'dashed', color = 'tomato', alpha = .5) +
+  # geom_segment(aes(x = iid_ci[1], xend = iid_ci[2], y = 1300, yend = 1300), color = 'turquoise3', alpha = .01) +
+  geom_vline(aes(xintercept = iid_ci[1]), linetype = 'dashed', color = 'turquoise3', alpha = .5) +
+  geom_vline(aes(xintercept = iid_ci[2]), linetype = 'dashed', color = 'turquoise3', alpha = .5) +
+  labs(fill = "Bootstrap Type", title='Bootstrapped Current Output', subtitle = 'Percentile confidence intervals given', 
+       x = 'Bootstrap Sample Means', y = 'Density') +
   theme_bw()
 ggsave("./images/mean_boot_plot.png", make_plot_for_pres(mean_boot_plot), width = 16, height = 12)
+```
+
+```{r}
+# rough MC coverage estimate assuming dependent bootstrap distribution as a null
+B <- 10000
+n <- nrow(boot_block_df)
+
+lower <- as.numeric(iid_ci[1])
+upper <- as.numeric(iid_ci[2])
+
+mu <- mean(boot_block_df$boot)
+sigma <- sd(boot_block_df$boot)
+
+sample <- rnorm(B, mu, sigma)
+
+sum(sample >= lower & sample <= upper) / n
 ```

### 2021-12-08 deliverables/presentation.Rmd
Added code for generating bootstrap histogram, and histogram in presentation
Reordered first slides

@@
 ---
+
+# Motivation
+
+- Renewable energy solutions
+- Problem: balance contribution of PV modules with demand on electrical grid
+  - Our work deals with the expected PV contribution
 
@@
   - Compare iid and time series bootstrap
-
-# Motivation
-
-- Understanding renewable energy sources like solar power is critical to the next century.
-- PV modules installed at individual households/nodes reduce the overall load on the electrical grid.
-- Balancing the expected contribution of an individual PV module with the expected residual demand required of the grid is a large optimization problem.
-    - We dealt with half of that optimization problem
 
@@
 
+![](./images/mean_boot_plot.png)
+
 # Future Research

### 2021-12-09 deliverables/paper.Rmd
Added Time Series Graphs
Added the graphs for blocks across different hour windows.

@@
 ```{r sampling}
+time_series_irradiance <- function(df, variable, var_name, hour_range) {
+  newdf <- df %>%
+    mutate(date = lubridate::make_datetime(Year, Month, Day, Hour, Minute)) %>%
+    # each day in this dataset is subset from 11:16, meaning there is 6 hours per day
+    head(hour_range * 6)
+
+  blocks <- c(newdf$date[seq(1,length(newdf$date), hour_range)])
+
+  re_plot <- ggplot(newdf) +
+    geom_line(aes_string(x = "date", y = variable)) +
+    geom_vline(xintercept = blocks, colour = "red") +
+    xlab("Date") +
+    ylab(var_name) +
+    theme_bw()
+  return(ret_plot)
+}
+
+sixhr_plot_current <- time_series_irradiance(df_high_irradiance, "current_out", "Current Output", 6)
+ggsave("./images/sixhr_plot_current.png", make_plot_for_pres(sixhr_plot_current))
+
+HOUR_WINDOW <- 8:19
+df_more_irradiance <- df[df$Hour %in% HOUR_WINDOW, ]
+filtered_current_hist <- ggplot() +
+  geom_histogram(aes(x = df_high_irradiance$current_out)) +
+  theme_bw()
+print(filtered_current_hist)
+
+twelvehr_plot_current <- time_series_irradiance(df_more_irradiance, "current_out", "Current Output", 12)
+ggsave("./images/twelvehr_plot_current.png", make_plot_for_pres(twelvehr_plot_current))
+
+day_plot_current <- time_series_irradiance(df, "current_out", "Current Output", 24)
+ggsave("./images/day_plot_current.png", make_plot_for_pres(day_plot_current))
+
 # B1 <- df[df$Hour == c(7:10),]$current_output # Hours before peak

### 2021-12-13 deliverables/paper.Rmd

@@
 
-  re_plot <- ggplot(newdf) +
+  ret_plot <- ggplot(newdf) +
     geom_line(aes_string(x = "date", y = variable)) +

### 2021-12-14 deliverables/paper.Rmd
Preprocessing section

@@
 ```
+
+# Preprocessing
+
+To translate raw irradiance data into meaningful current data, several steps had to be taken to simulate the electrical workings of a real PV module.
+First, adjustments had to be made to account for the fact that an imperfect amount of irradiance actually reaches a PV module for capture.
+Then, effective irradiance must be translated into output current and maximized with an attached power processor.
+The process and accompanying assumptions are given below.
+
+## Effective Irradiance
+
+The effective irradiance that reaches a PV module is different from the total irradiance emitted by the sun due to a variety of small considerations that must be made.
+Among these are the tilt and rotation of the Earth, the latitude and longitude of the PV module, and the inclination of the panel towards the horizon.
+Most simply, $I_{eff} \propto I_{sun} cos \theta(t, N)$ where $\theta \in [-\frac{\pi}{2},\frac{\pi}{2}]$ is the angle between the sunlight and the surface of the solar cell.
+$$
+\begin{aligned}
+cos \theta(t, N) = & sin \gamma(N) sin La cos \beta - \\
+& sin \gamma(N) cos La sin \beta cos \alpha + \\
+& cos \gamma(N) cos La cos \beta cos w(t,N) + \\
+& cos \gamma(N) sin La sin \beta cos \alpha cos w(t,N) + \\
+& cos \gamma(N) sin \beta sin \alpha sin w(t,N) \\
+\end{aligned}
+$$
+Here, $\gamma (N) = sin^{-1}[sin(.409)sin(D(N))]$ is known as the declination angle and is calculated from the angle of the Earth's axis and a metric $D(N) = 360(N - 81)/360$ where $N \in [1, 365]$ is the day of the year.
+$La \in [0, \frac{\pi}{2}]$ is the latitude of the PV module in radians.
+$\alpha = .524$ corresponds to whether the panel faces East or West and $\beta = .785$ describes the angle between the panel and the horizon; these values were supplied as assumptions.
+Finally, the hour angle $w(t, N)$ accounts for the angle of sunlight given the rotation of the Earth.
+$$
+\begin{aligned}
+w(t,N) &= 15 (AST(t,N) - 12) \\
+AST(t,N) &= t + \Delta t + ET(N) \\
+&= t + \frac{Lo - 15 UTC_{off}}{15} + ET(N) \\
+ET(N) &= [9.87sin(2D(N)) - 7.53cos(D(N)) - 1.5sin(D(N))] / 60
+\end{aligned}
+$$
+Here, $AST(t,N) \in [0,24]$ is the apparent solar time relative to the sun and is calculated as an adjustment to the current time $t$ based on longitude $Lo$, an offset from the coordinated universal time $UTC_{off}$, and the equation of time $ET(N)$.
+The above calculations produce results in degrees but values were switched to radians before trigonometric functions were applied.
+Ultimately, $I_{eff} \propto I_{sun} cos \theta(t, N)$ resulted from our astronomical adjustments and was used for PV module calculations.
+
+## PV Module
+
+Within the PV module, effective irradiance is translated into output current but the entire process is regulated by an attached power processor that ensure that the maximum power is extracted.
+It does this by testing a range of open circuit voltages $v_{oc}$ between zero and an assumed maximum $v_{oc} = 1.8$ on the PV module's I-V curve which describes how current and voltage vary together for each solar cell in the module.
+Here, the I-V curve is given by
+$$
+i_{out} = i_{l} - i_{o}[e^{\frac{qv}{n \kappa T}} - 1]
+$$ 
+with constants $q = 1.6*10^{-19}$ and $\kappa = 1.38 * 10^{-23}$.
+The light-generated current $i_{l}(t,N) = .0005 I_{eff}(t,N)$ which adjusts $I_{eff}$ for the assumed short circuit current $i_{sc} = 5$ and a radiation rate factor.
+With a leakage factor $i_{o}$ and the ambient temperature $T$ in degrees Kelvin, $i_{out}$ can be calculated for various voltage values.
+The combined power of the module is then calculated by
+$$
+P = n_{p} n_{s} max\{i_{out} v\}
+$$ 
+where $n_{p} = 6$ is the assumed number of parallel-wired solar cells in the module and $n_{s} = 6$ is the assumed number of series-wired solar cells in the module.
+A final conversion efficiency factor $\eta = .211$ is applied and $P'_{max} = \eta P_{max}$ is returned.
 

### 2021-12-14 deliverables/paper.Rmd
Added references, reorganized code, reorganized paper, fixed preprocessing CSV

--- 2021-12-13 deliverables/paper.Rmd
@@
 output: pdf_document
+bibliography: references.bib
 ---
 
-```{r reprod}
+```{r reprod, include=FALSE}
+# Reproducibility
 set.seed(400)
@@
 ```{r libs, include=FALSE, message=FALSE}
+# Import packages
+library(knitr)
 library(tidyverse)
@@
 
-```{r data, message=FALSE, warning=FALSE}
-df <- read_csv("../data/final/irradiance_full_final.csv") %>%
-  select(-'...1')
+```{r data, include=FALSE, message=FALSE, warning=FALSE}
+# Read in preprocessed data
+df <- read_csv("../data/final/irradiance_full_final.csv")
 ```
 
-```{r plot-pres-func}
+```{r plot-pres-func, include=FALSE}
+# Function to format plots for use in presentation
 make_plot_for_pres <- function(plot) {
@@
 
+# Background
+
+TODO
+
+# Data
+
+The data used was from the National Solar Radiation Database (NSRDB),
+provided by the National Renewable Energy Laboratory
+(NREL) (@nsrdb).
+The data was acquired via the NSRDB API (@nsrdb_api).
+
+The model paper for this project, @miozzo_2014
+used data cited from the NSRDB.
+However,
+there were some discrepancies between our data
+and the data used in @miozzo_2014.
+Firstly, the year range in the paper was 1991--2010,
+while the year range available when we acquired
+the data from the NSRDB API was 1998--2010.
+Secondly, the primary variable used in 
+was extraterrestrial radiation,
+"the amount of global horizontal radiation that a location on Earth would
+receive if there was no atmosphere" (@nrel_glossary).
+Our primary measure was global horizontal radiation (GHI),
+which is the total solar radiation to reach
+a horizontal plane on the Earth's surface (@nrel_glossary).
+Because ETR and GHI are closely related,
+we felt that the data was acceptable for a
+replication of the process in @miozzo_2014,
+with the caveat that the calculations may only be approximate.
+
+## Data filtering
+
+```{r filter, include=FALSE}
+HOUR_WINDOW <- 11:16
+df_high_irradiance <- df[df$Hour %in% HOUR_WINDOW, ]
+filtered_current_hist <- ggplot() +
+  geom_histogram(aes(x = df_high_irradiance$current_out, y = ..density..), binwidth = .0025) +
+  theme_bw() +
+  labs(
+    title = "Distribution of Current Outputs", subtitle = "Restricted to high irradiance hours of the day",
+    x = "Current Output", y = "Density"
+  )
+print(filtered_current_hist)
+```
+
 # EDA
 
-```{r timeseries, warning=FALSE}
+```{r timeseries, include=FALSE, warning=FALSE}
 
@@
 
-```{r hour-plot}
+```{r hour-plot, include=FALSE}
 
@@
 
-# Data filtering
+# Preprocessing
 
-```{r filter}
-HOUR_WINDOW <- 11:16
-df_high_irradiance <- df[df$Hour %in% HOUR_WINDOW, ]
-filtered_current_hist <- ggplot() +
-  geom_histogram(aes(x = df_high_irradiance$current_out, y = ..density..), binwidth = .0025) +
-  theme_bw() +
-  labs(title='Distribution of Current Outputs', subtitle='Restricted to high irradiance hours of the day', 
-       x = 'Current Output', y = 'Density')
-print(filtered_current_hist)
-```
+To translate raw irradiance data into meaningful current data, several steps
+had to be taken to simulate the electrical workings of a real PV module. First,
+adjustments had to be made to account for the fact that an imperfect amount of
+irradiance actually reaches a PV module for capture. Then, effective irradiance
+must be translated into output current and maximized with an attached power
+processor. The process and accompanying assumptions are given below.
+
+## Effective Irradiance
+
+The effective irradiance that reaches a PV module is different from the total
+irradiance emitted by the sun due to a variety of small considerations that
+must be made. Among these are the tilt and rotation of the Earth, the latitude
+and longitude of the PV module, and the inclination of the panel towards the
+horizon. Most simply, $I_{eff} \propto I_{sun} cos \theta(t, N)$ where $\theta
+\in [-\frac{\pi}{2},\frac{\pi}{2}]$ is the angle between the sunlight and the
+surface of the solar cell [@miozzo_2014, p. 689]. 
+$$
+\begin{aligned}
+cos \theta(t, N) = & sin \gamma(N) sin La cos \beta - \\
+& sin \gamma(N) cos La sin \beta cos \alpha + \\
+& cos \gamma(N) cos La cos \beta cos \omega(t,N) + \\
+& cos \gamma(N) sin La sin \beta cos \alpha cos \omega(t,N) + \\
+& cos \gamma(N) sin \beta sin \alpha sin \omega(t,N) \\
+\end{aligned}
+$$
+Here, $\gamma (N) = sin^{-1}[sin(.409)sin(D(N))]$ is known as the declination angle and is calculated from the angle of the Earth's axis and a metric $D(N) = 360(N - 81)/360$ where $N \in [1, 365]$ is the day of the year [@miozzo_2014, p.
+689].
+$La \in [0, \frac{\pi}{2}]$ is the latitude of the PV module in radians [@miozzo_2014, p.
+689].
+$\alpha = .524$ corresponds to whether the panel faces East or West and $\beta = .785$ describes the angle between the panel and the horizon [@miozzo_2014, p.
+689].
+The constants $\alpha$ and $\beta$
+were supplied as assumptions based on the location of the NSRDB
+data [@miozzo_2014, p. 691].
+Finally, the hour angle $\omega(t, N)$ accounts for the angle of sunlight given the rotation of the Earth [@miozzo_2014, p. 689]. 
+$$
+\begin{aligned}
+w(t,N) &= 15 (AST(t,N) - 12) \\
+AST(t,N) &= t + \Delta t + ET(N) \\
+&= t + \frac{Lo - 15 UTC_{off}}{15} + ET(N) \\
+ET(N) &= [9.87sin(2D(N)) - 7.53cos(D(N)) - 1.5sin(D(N))] / 60
+\end{aligned}
+$$
+Here, $AST(t,N) \in [0,24]$ is the apparent solar time relative to the sun and is calculated as an adjustment to the current time $t$ based on longitude $Lo$, an offset from the coordinated universal time $UTC_{off}$, and the equation of time $ET(N)$ [@miozzo_2014, p.
+689].
+The above calculations produce results in degrees but values were switched to radians before trigonometric functions were applied.
+Ultimately, $I_{eff} \propto I_{sun} cos \theta(t, N)$ resulted from our astronomical adjustments and was used for PV module
+calculations [@miozzo_2014, p. 689].
+
+## PV Module
+
+Within the PV module, effective irradiance is translated into output current
+but the entire process is regulated by an attached power processor that ensure
+that the maximum power is extracted. It does this by testing a range of open
+circuit voltages $v_{oc}$ between zero and an assumed maximum $v_{oc} = 1.8$ on
+the PV module's I-V curve which describes how current and voltage vary together
+for each solar cell in the module [@miozzo_2014, p. 691].
+Here, the I-V curve is given by 
+$$
+i_{out} = i_{l} - i_{o}[e^{\frac{qv}{n \kappa T}} - 1]
+$$ 
+with constants $q = 1.6*10^{-19}$ and $\kappa = 1.38 * 10^{-23}$.
+The light-generated current $i_{l}(t,N) = .0005 I_{eff}(t,N)$ which adjusts $I_{eff}$ for the assumed short circuit current $i_{sc} = 5$ and a radiation rate factor [@miozzo_2014, p.
+691].
+With a leakage factor $i_{o}$ and the ambient temperature $T$ in degrees Kelvin, $i_{out}$ can be calculated for various voltage values. The combined power of the module is then calculated by 
+$$
+P = n_{p} n_{s} max\{i_{out} v\}
+$$ 
+where $n_{p} = 6$ is the assumed number of parallel-wired solar cells in the module and $n_{s} = 6$ is the assumed number of series-wired solar cells in the module [@miozzo_2014, p. 691].
+A final conversion efficiency factor $\eta = .211$ is applied and $P'_{max} = \eta P_{max}$ is returned.
 
@@
 
-```{r functions}
+```{r functions, include=FALSE}
 
@@
 mean_boot_plot <- ggplot(vis_df) +
-  geom_histogram(aes(x = boot, y = ..density.., fill = type), position = "identity", alpha = .5, color = "black", 
-                 binwidth = .0001) +
+  geom_histogram(aes(x = boot, y = ..density.., fill = type),
+    position = "identity", alpha = .5, color = "black",
+    binwidth = .0001
+  ) +
   # geom_segment(aes(x = block_ci[1], xend = block_ci[2], y = 1250, yend = 1250), color = 'tomato', alpha = .01) +
-  geom_vline(aes(xintercept = block_ci[1]), linetype = 'dashed', color = 'tomato', alpha = .5) +
-  geom_vline(aes(xintercept = block_ci[2]), linetype = 'dashed', color = 'tomato', alpha = .5) +
+  geom_vline(aes(xintercept = block_ci[1]), linetype = "dashed", color = "tomato", alpha = .5) +
+  geom_vline(aes(xintercept = block_ci[2]), linetype = "dashed", color = "tomato", alpha = .5) +
   # geom_segment(aes(x = iid_ci[1], xend = iid_ci[2], y = 1300, yend = 1300), color = 'turquoise3', alpha = .01) +
-  geom_vline(aes(xintercept = iid_ci[1]), linetype = 'dashed', color = 'turquoise3', alpha = .5) +
-  geom_vline(aes(xintercept = iid_ci[2]), linetype = 'dashed', color = 'turquoise3', alpha = .5) +
-  labs(fill = "Bootstrap Type", title='Bootstrapped Current Output', subtitle = 'Percentile confidence intervals given', 
-       x = 'Bootstrap Sample Means', y = 'Density') +
+  geom_vline(aes(xintercept = iid_ci[1]), linetype = "dashed", color = "turquoise3", alpha = .5) +
+  geom_vline(aes(xintercept = iid_ci[2]), linetype = "dashed", color = "turquoise3", alpha = .5) +
+  labs(
+    fill = "Bootstrap Type", title = "Bootstrapped Current Output", subtitle = "Percentile confidence intervals given",
+    x = "Bootstrap Sample Means", y = "Density"
+  ) +
   theme_bw()
@@
 
-```{r}
+```{r, include=FALSE}
 # rough MC coverage estimate assuming dependent bootstrap distribution as a null
@@
 
-```{r sampling}
+```{r sampling, include=FALSE}
 time_series_irradiance <- function(df, variable, var_name, hour_range) {
@@
 
-  blocks <- c(newdf$date[seq(1,length(newdf$date), hour_range)])
+  blocks <- c(newdf$date[seq(1, length(newdf$date), hour_range)])
 
@@
 
+# Future Work
+
+For future work, it would be good to examine the wider implications
+of the inferential results explored in this paper.
+The model paper for this project simulated results for small solar panels
+in a specific configuration,
+making it hard to generalize the results to any meaningful solar array.
+However, if the results could be generalized,
+it would be good to see how current output based on solar irradiance
+fits into a larger model of uncertainty surrounding renewable energy generation.
+One example of such an approach would be @constante_2018,
+in which PV output simulated from solar irradiance
+is used in conjunction with other sources of of uncertainty.
+The estimates of PV output in this paper are only part
+of the story with respect to uncertainty in PV power generation.
+Other relevant types of uncertainty are uncertainty of energy demand,
+uncertainty in personal PV module placement on the grid,
+and uncertainty in the power generation of other renewable energy
+sources, to name a few (@constante_2018).
+
+\newpage
+
+# References
+
+<div id="refs"></div>
+
+\newpage
+
+# Appendix
+
+```{r show-code, ref.label = all_labels(), echo = TRUE, eval = FALSE}
+
+```
+
 - We chose our initial paper because of its use of Monte Carlo.

### 2021-12-14 data_preprocessing/main.py

@@ -46,7 +46,7 @@ def main():
         axis=1,
     )
 
-    df.to_csv(PROCESSED_CSV)
+    df.to_csv(PROCESSED_CSV, index=False)
 
 if __name__ == "__main__":
     main()

### 2021-12-14 deliverables/paper.Rmd
Added initial methods section with bootstrapping algorithm and justification

@@
 
-# Bootstrapping
+# Methods
 
@@
 
+The main goal of this paper is to provide point and interval
+estimates for average (simulated) current output
+based on solar irradiance.
+Given that the sample size is large,
+a simple average may have been sufficient.
+However, we were interested in characterizing
+the difference in performance between
+standard bootstrapping methods and dependent bootstrapping methods.
+The current output provides a good test case for this.
+Both the preprocessed and postprocessed currents are not independent data;
+values of either irradiance or current output
+depend on surrounding values in time.
+As such, standard methods for bootstrapping should be
+less effective at estimating the true mean irradiance output.
+
+The standard (*iid*) bootstrap procedure begins with
+the assumption that our original sample consists of
+independent and identically distributed draws from a distribution.
+The univariate bootstrap then provides a distribution of bootstrap
+estimates for a statistic,
+$\hat{\theta}^{(1)}, \ldots, \{\hat{\theta}^{(B)}$,
+where $B$ is the number of bootstrap replicates
+and $\theta$ is a statistic of interest.
+For our purposes, $\theta$ is simply the average current output
+for our specific solar module configuration.
+The *iid* bootstrap constructs the distribution of bootstrap estimates
+as follows:
+
+For $i = 1, \ldots, B$:
+
+1. Sample $n$ values $x_j$ from the data with replacement
+1. Calculate
+   $\hat{\theta}^{(i)} = \frac{1}{n} \sum_{j = 1}^{n} x_j$
+   for the bootstrap sample.
+
+In contrast, a dependent bootstrap does not assume
+the original sample was independent and identically distributed.
+To capture the dependence in the data,
+we chose to use a nonparametric non-overlapping block bootstrap.
+This is because during the data filtering step,
+we excluded nighttime values from the dataset.
+This meant we essentially had predefined,
+mostly independent blocks from which to sample:
+the daytime current values for each day.
+Given that our daytime current values for each day
+could be used as non-overlapping blocks,
+our dataset could be seen as a collection
+of $n_B$ blocks.
+
+For i = $1, \ldots, B$:
+
+1. Sample $n_B$ blocks (days)
+$\mathcal{B}_{1}, ..., \mathcal{B}_{n_B}$
+from the data with replacement.
+1. Calculate
+   $\hat{\theta}^{(i)} = \frac{1}{n} \sum_{j = 1}^{n} x_j$
+   on a new dataset
+  $\mathcal{B}_{1} \cup \cdots \cup \mathcal{B}_{n_B}$.
+
+The new dataset in Step 2
+corresponds to sampling and rearranging days to make
+a new dataset the same size as the original.
+This procedure preserves the dependence within days,
+while still allowing resampling.
+
+# Results
+
 # Future Work

### 2021-12-15 deliverables/paper.Rmd
Confidence interval paragraph

@@
 
+A percentile bootstrap confidence interval was calculated for both the dependent and *iid* estimates.
+This method was chosen for its simplicity and for the fact that the bootstrapped sampling distributions agreed with the assumption of normality under central limit theorem.
+Performing an *iid* bootstrap on dependent data should yield an artificially narrow confidence interval and this belief was supported by the relative widths of the resulting intervals.
+
 # Future Work

### 2021-12-15 deliverables/paper.Rmd
Added figures and tables, reorganized

@@
 date: "2021/12/16"
-output: pdf_document
+output: bookdown::pdf_document2
 bibliography: references.bib
@@
 library(knitr)
+library(bookdown)
 library(tidyverse)
 library(boot)
+```
+
+```{r knitr-opts}
+knitr::opts_chunk$set(fig.pos = "t", fig.align = 'center', out.width = "75%", out.extra = "")
 ```
@@
 with the caveat that the calculations may only be approximate.
-
-## Data filtering
-
-```{r filter, include=FALSE}
-HOUR_WINDOW <- 11:16
-df_high_irradiance <- df[df$Hour %in% HOUR_WINDOW, ]
-filtered_current_hist <- ggplot() +
-  geom_histogram(aes(x = df_high_irradiance$current_out, y = ..density..), binwidth = .0025) +
-  theme_bw() +
-  labs(
-    title = "Distribution of Current Outputs", subtitle = "Restricted to high irradiance hours of the day",
-    x = "Current Output", y = "Density"
-  )
-print(filtered_current_hist)
-```
-
-# EDA
-
-```{r timeseries, include=FALSE, warning=FALSE}
-
-#' Plot a timeseries of a single variable
-#'
-#' @param df A dataframe containing `variable`
-#' @param variable A string referring to a column of `df`
-#' @param var_name The human-readable name of `variable`, for labelling
-#' @param n_days The number of days (starting from the
-#'    beginning of the data) to plot
-#' @return A ggplot object of a timeseries of `variable` for `n_days` days
-time_series_of_var <- function(df, variable, var_name, n_days) {
-  ret_plot <- df %>%
-    mutate(date = lubridate::make_datetime(Year, Month, Day, Hour, Minute)) %>%
-    head(24 * n_days) %>%
-    ggplot() +
-    geom_line(aes_string(x = "date", y = variable)) +
-    xlab("Date") +
-    ylab(var_name) +
-    theme_bw()
-  return(ret_plot)
-}
-
-time_series_ghi <- time_series_of_var(df, "GHI", "GHI", 6)
-time_series_current <- time_series_of_var(df, "current_out", "Output current", 6)
-ggsave("./images/time_series_ghi.png", make_plot_for_pres(time_series_ghi))
-ggsave("./images/time_series_current.png", make_plot_for_pres(time_series_current))
-```
-
-```{r hour-plot, include=FALSE}
-
-#' Plot the values of `variable` at each hour of the day
-#'
-#' @param df A dataframe containing `variable`
-#' @param variable A string referring to a column of `df`
-#' @param var_name The human-readable name of `variable`, for labelling
-#' @return A ggplot object of a boxplot of `variable` segmented by hour
-hour_plot_of_var <- function(df, variable, var_name) {
-  df$Hour <- as.factor(df$Hour)
-  ret_plot <- df %>%
-    ggplot() +
-    geom_boxplot(aes_string(x = "Hour", y = variable)) +
-    xlab("Hour (0-23)") +
-    ylab(paste("Distribution of", var_name)) +
-    scale_x_discrete(breaks = seq(0, 23, 2)) +
-    theme_bw()
-  return(ret_plot)
-}
-
-hour_plot_ghi <- hour_plot_of_var(df, "GHI", "GHI")
-hour_plot_current <- hour_plot_of_var(df, "current_out", "output current")
-ggsave("./images/hour_plot_ghi.png", make_plot_for_pres(hour_plot_ghi))
-ggsave("./images/hour_plot_current.png", make_plot_for_pres(hour_plot_current))
-```
 
@@
 # Methods
+
+## Data filtering
+
+```{r filter, include=FALSE}
+HOUR_WINDOW <- 11:16
+df_high_irradiance <- df[df$Hour %in% HOUR_WINDOW, ]
+filtered_current_hist <- ggplot() +
+  geom_histogram(aes(x = df_high_irradiance$current_out, y = ..density..), binwidth = .0025) +
+  theme_bw() +
+  labs(
+    title = "Distribution of Current Outputs", subtitle = "Restricted to high irradiance hours of the day",
+    x = "Current Output", y = "Density"
+  )
+print(filtered_current_hist)
+```
+
+## Bootstrapping
+
+The main goal of this paper is to provide point and interval
+estimates for average (simulated) current output
+based on solar irradiance.
+Given that the sample size is large,
+a simple average may have been sufficient.
+However, we were interested in characterizing
+the difference in performance between
+standard bootstrapping methods and dependent bootstrapping methods.
+The current output provides a good test case for this.
+Both the preprocessed and postprocessed currents are not independent data;
+values of either irradiance or current output
+depend on surrounding values in time.
+As such, standard methods for bootstrapping should be
+less effective at estimating the true mean irradiance output.
+
+The standard (*iid*) bootstrap procedure begins with
+the assumption that our original sample consists of
+independent and identically distributed draws from a distribution.
+The univariate bootstrap then provides a distribution of bootstrap
+estimates for a statistic,
+$\hat{\theta}^{(1)}, \ldots, \{\hat{\theta}^{(B)}$,
+where $B$ is the number of bootstrap replicates
+and $\theta$ is a statistic of interest.
+For our purposes, $\theta$ is simply the average current output
+for our specific solar module configuration.
+The *iid* bootstrap constructs the distribution of bootstrap estimates
+as follows:
+
+For $i = 1, \ldots, B$:
+
+1. Sample $n$ values $x_j$ from the data with replacement
+1. Calculate
+   $\hat{\theta}^{(i)} = \frac{1}{n} \sum_{j = 1}^{n} x_j$
+   for the bootstrap sample.
+
+In contrast, a dependent bootstrap does not assume
+the original sample was independent and identically distributed.
+To capture the dependence in the data,
+we chose to use a nonparametric non-overlapping block bootstrap.
+This is because during the data filtering step,
+we excluded nighttime values from the dataset.
+This meant we essentially had predefined,
+mostly independent blocks from which to sample:
+the daytime current values for each day.
+Given that our daytime current values for each day
+could be used as non-overlapping blocks,
+our dataset could be seen as a collection
+of $n_B$ blocks.
+
+For i = $1, \ldots, B$:
+
+1. Sample $n_B$ blocks (days)
+$\mathcal{B}_{1}, ..., \mathcal{B}_{n_B}$
+from the data with replacement.
+1. Calculate
+   $\hat{\theta}^{(i)} = \frac{1}{n} \sum_{j = 1}^{n} x_j$
+   on a new dataset
+  $\mathcal{B}_{1} \cup \cdots \cup \mathcal{B}_{n_B}$.
+
+The new dataset in Step 2
+corresponds to sampling and rearranging days to make
+a new dataset the same size as the original.
+This procedure preserves the dependence within days,
+while still allowing resampling.
+
+A percentile bootstrap confidence interval was calculated for both the
+dependent and *iid* estimates. This method was chosen for its simplicity and
+for the fact that the bootstrapped sampling distributions agreed with the
+assumption of normality under central limit theorem. Performing an *iid*
+bootstrap on dependent data should yield an artificially narrow confidence
+interval and this belief was supported by the relative widths of the resulting
+intervals.
+
+# Results
+
+## Descriptive Statistics
+
+```{r timeseries, include=FALSE, warning=FALSE}
+
+#' Plot a timeseries of a single variable
+#'
+#' @param df A dataframe containing `variable`
+#' @param variable A string referring to a column of `df`
+#' @param var_name The human-readable name of `variable`, for labelling
+#' @param n_days The number of days (starting from the
+#'    beginning of the data) to plot
+#' @return A ggplot object of a timeseries of `variable` for `n_days` days
+time_series_of_var <- function(df, variable, var_name, n_days) {
+  ret_plot <- df %>%
+    mutate(date = lubridate::make_datetime(Year, Month, Day, Hour, Minute)) %>%
+    head(24 * n_days) %>%
+    ggplot() +
+    geom_line(aes_string(x = "date", y = variable)) +
+    xlab("Date") +
+    ylab(var_name) +
+    theme_bw()
+  return(ret_plot)
+}
+
+time_series_ghi <- time_series_of_var(df, "GHI", "GHI", 6)
+time_series_current <- time_series_of_var(df, "current_out", "Output current", 6)
+ggsave("./images/time_series_ghi.png", make_plot_for_pres(time_series_ghi))
+ggsave("./images/time_series_current.png", make_plot_for_pres(time_series_current))
+```
+
+```{r hour-plot, include=FALSE}
+
+#' Plot the values of `variable` at each hour of the day
+#'
+#' @param df A dataframe containing `variable`
+#' @param variable A string referring to a column of `df`
+#' @param var_name The human-readable name of `variable`, for labelling
+#' @return A ggplot object of a boxplot of `variable` segmented by hour
+hour_plot_of_var <- function(df, variable, var_name) {
+  df$Hour <- as.factor(df$Hour)
+  ret_plot <- df %>%
+    ggplot() +
+    geom_boxplot(aes_string(x = "Hour", y = variable)) +
+    xlab("Hour (0-23)") +
+    ylab(paste("Distribution of", var_name)) +
+    scale_x_discrete(breaks = seq(0, 23, 2)) +
+    theme_bw()
+  return(ret_plot)
+}
+
+hour_plot_ghi <- hour_plot_of_var(df, "GHI", "GHI")
+hour_plot_current <- hour_plot_of_var(df, "current_out", "output current")
+ggsave("./images/hour_plot_ghi.png", make_plot_for_pres(hour_plot_ghi))
+ggsave("./images/hour_plot_current.png", make_plot_for_pres(hour_plot_current))
+```
+
+```{r ghi-ts, fig.cap="Time series of GHI values for the first six days in the dataset", echo=FALSE}
+print(time_series_ghi)
+```
+
+```{r current-ts, fig.cap="Time series of simulated current values for the first six days in the dataset.", echo=FALSE}
+print(time_series_current)
+```
+
+```{r ghi-hour-boxplot, fig.cap="Distribution of GHI at each hour for the entire unfiltered dataset.", echo=FALSE}
+print(hour_plot_ghi)
+```
+
+```{r current-hour-boxplot, fig.cap="Distribution of GHI at each hour for the entire unfiltered dataset.", echo=FALSE}
+print(hour_plot_current)
+```
+
+## Bootstrapping
 
@@
 
-```{r, include=FALSE}
-# rough MC coverage estimate assuming dependent bootstrap distribution as a null
-B <- 10000
-n <- nrow(boot_block_df)
+```{r mean-boot-plot, fig.cap="Distributions of bootstrap mean simulated current for both iid and block bootstrapping methods.
+Percentile 95 percent bootstrap confidence intervals for mean simulated current overlayed.", echo=FALSE}
+print(mean_boot_plot)
+```
 
-lower <- as.numeric(iid_ci[1])
-upper <- as.numeric(iid_ci[2])
+```{r bootstrap-ci, echo=FALSE}
+bootstrap_ci_table <- rbind(
+c("iid", paste("(", round(iid_ci[1],3),",", round(iid_ci[2],3), ")")),
+c("Block", paste("(", round(block_ci[1],3),",", round(block_ci[2], 3), ")"))
+)
+colnames(bootstrap_ci_table) <- c("Bootstrap Method", "95% CI")
+rownames(bootstrap_ci_table) <- c()
 
-mu <- mean(boot_block_df$boot)
-sigma <- sd(boot_block_df$boot)
+bootstrap_ci_table %>% kable(booktabs = TRUE,
+caption="95% confidence intervals for the mean simulated current, as generated by each of the two bootstrapping methods.")
 
-sample <- rnorm(B, mu, sigma)
-
-sum(sample >= lower & sample <= upper) / n
 ```
@@
 
-The main goal of this paper is to provide point and interval
-estimates for average (simulated) current output
-based on solar irradiance.
-Given that the sample size is large,
-a simple average may have been sufficient.
-However, we were interested in characterizing
-the difference in performance between
-standard bootstrapping methods and dependent bootstrapping methods.
-The current output provides a good test case for this.
-Both the preprocessed and postprocessed currents are not independent data;
-values of either irradiance or current output
-depend on surrounding values in time.
-As such, standard methods for bootstrapping should be
-less effective at estimating the true mean irradiance output.
-
-The standard (*iid*) bootstrap procedure begins with
-the assumption that our original sample consists of
-independent and identically distributed draws from a distribution.
-The univariate bootstrap then provides a distribution of bootstrap
-estimates for a statistic,
-$\hat{\theta}^{(1)}, \ldots, \{\hat{\theta}^{(B)}$,
-where $B$ is the number of bootstrap replicates
-and $\theta$ is a statistic of interest.
-For our purposes, $\theta$ is simply the average current output
-for our specific solar module configuration.
-The *iid* bootstrap constructs the distribution of bootstrap estimates
-as follows:
-
-For $i = 1, \ldots, B$:
-
-1. Sample $n$ values $x_j$ from the data with replacement
-1. Calculate
-   $\hat{\theta}^{(i)} = \frac{1}{n} \sum_{j = 1}^{n} x_j$
-   for the bootstrap sample.
-
-In contrast, a dependent bootstrap does not assume
-the original sample was independent and identically distributed.
-To capture the dependence in the data,
-we chose to use a nonparametric non-overlapping block bootstrap.
-This is because during the data filtering step,
-we excluded nighttime values from the dataset.
-This meant we essentially had predefined,
-mostly independent blocks from which to sample:
-the daytime current values for each day.
-Given that our daytime current values for each day
-could be used as non-overlapping blocks,
-our dataset could be seen as a collection
-of $n_B$ blocks.
-
-For i = $1, \ldots, B$:
-
-1. Sample $n_B$ blocks (days)
-$\mathcal{B}_{1}, ..., \mathcal{B}_{n_B}$
-from the data with replacement.
-1. Calculate
-   $\hat{\theta}^{(i)} = \frac{1}{n} \sum_{j = 1}^{n} x_j$
-   on a new dataset
-  $\mathcal{B}_{1} \cup \cdots \cup \mathcal{B}_{n_B}$.
-
-The new dataset in Step 2
-corresponds to sampling and rearranging days to make
-a new dataset the same size as the original.
-This procedure preserves the dependence within days,
-while still allowing resampling.
-
-# Results
-
-A percentile bootstrap confidence interval was calculated for both the dependent and *iid* estimates.
-This method was chosen for its simplicity and for the fact that the bootstrapped sampling distributions agreed with the assumption of normality under central limit theorem.
-Performing an *iid* bootstrap on dependent data should yield an artificially narrow confidence interval and this belief was supported by the relative widths of the resulting intervals.
-
 # Future Work

### 2021-12-15 deliverables/paper.Rmd
Introduction
I added an introduction to the paper and related our current output/peak hour bootstrapping to the total compensation of installing solar panels. A lot of articles online discussed whether or not it is worth it for a reader to install solar panels in their area. Figured we could use our output to determine whether it is worth it for Los Angeles residents.

@@
 
-TODO
+This project began with the search for a research paper
+utilizing the process of Monte Carlo sampling or other 
+statistcal computing methods. We landed on a few papers 
+that did exatcly this, one titled, "SolarStat: Modeling 
+Photovoltaic Sources through Stochastic Markov Processes". 
+Instead of comparing our works of Monte Carlo simulations 
+we were able to utilize a more recent topic discussed in class.
+This of course was bootsrapping, a topic we didn't get too in
+depth about; however, everything needed to know will be explained 
+later on in the paper. Regarding our choice of articles, it was a
+question of interest and importance. During the last decade the world
+has seen a massive uptick in renewable energy pushes. With this 
+new wave of renewables and the rather large misunderstanding of their
+total power we decided it would be interesting to look into solar and 
+what it was all about. The big question being asked regarding
+solar power is whether or not it is worth heading the initial cost
+of set up and general mantienance. Many websites inform users of 
+just this, according to the Solar.com Learning Center, the average
+solar panel net cost is $2,700 per 1.5 kilowatt-hour system (@solar.com). 
+The biggest upside to solar panels is being able to slice down your 
+energy bill, and eventually have it pay for itself. In order to 
+calculate a solar panels net output experts recommend looking at three things. 
+1. Determine your locations peak sun hours (average US around 4 hours/day)
+2. Determine your energy compensation (average $0.03/kilowatt-hour (kWh))
+3. Determine your solar panels net energy output
+Once you have calculated these three numbers it is as easy as 
+multiplying them together to receive your daily solar panel 
+compensation (@energysage). As an example Energy Sage looked at 
+California receiving 5 peak hours of sunlight a day with a single 
+solar panel producing 290 watts/hour. Daily, this panel will produce 
+around 1.5 kWh per day yielding the owner a rough compensation 
+of $16.42 a year.
+
+This paper looks at data from a 12 year period to determine the current 
+output of a specified city. It can be used to determine 
+on a month by month basis if a location produces a sufficient amount 
+of current from a solar panel. Providing customers with an idea on 
+whether or not they should make the investment that is solar.
 

### 2021-12-15 data_preprocessing/main.py
Changed data preprocessing to use power and not current

@@ -7,7 +7,7 @@ from pv_module import PVModule as pvm
 
 def main():
     SOLAR_IRRADIANCE = "irradiance_eff"
-    CURRENT_OUTPUT = "current_out"
+    POWER_OUTPUT = "power_out"
     RAW_CSV = "../data/raw/irradiance_full_raw.csv"
     PROCESSED_CSV =  "../data/final/irradiance_full_final.csv"
 
@@ -32,11 +32,11 @@ def main():
         axis=1,
     )
 
-    df[CURRENT_OUTPUT] = df.apply(
+    df[POWER_OUTPUT] = df.apply(
         lambda row: pp.estimated_power(
             ideality=1.5,
             temperature=row["Temperature"],
-            sc_current_module=5,
+            sc_power_module=5,
             i_eff=row[SOLAR_IRRADIANCE],
             n_parallel=6,
             n_series=6,

### 2021-12-15 data_preprocessing/power_processor.py

@@ -56,9 +56,8 @@ class PowerProcessor:
         vec_to_be_maximized = (
             n_parallel * n_series * np.multiply(current_vec, voltage_vec)
         )
-        max_index = np.argmax(vec_to_be_maximized)
-        # Return the *current* corresponding to the max power
-        return current_vec.item(max_index)
+        max_power = np.amax(vec_to_be_maximized)
+        return max_power
 
     @classmethod
     def estimated_power(

### 2021-12-15 data_preprocessing/main.py
Fixed unnecessary word replacement

@@ -36,7 +36,7 @@ def main():
         lambda row: pp.estimated_power(
             ideality=1.5,
             temperature=row["Temperature"],
-            sc_power_module=5,
+            sc_current_module=5,
             i_eff=row[SOLAR_IRRADIANCE],
             n_parallel=6,
             n_series=6,

### 2021-12-15 deliverables/paper.Rmd
Removed bindwith for new scale

@@
 just this, according to the Solar.com Learning Center, the average
-solar panel net cost is $2,700 per 1.5 kilowatt-hour system (@solar.com). 
+solar panel net cost is \$2,700 per 1.5 kilowatt-hour system (@solar.com). 
 The biggest upside to solar panels is being able to slice down your 
@@
 1. Determine your locations peak sun hours (average US around 4 hours/day)
-2. Determine your energy compensation (average $0.03/kilowatt-hour (kWh))
+2. Determine your energy compensation (average \$0.03/kilowatt-hour (kWh))
 3. Determine your solar panels net energy output
@@
 around 1.5 kWh per day yielding the owner a rough compensation 
-of $16.42 a year.
+of \$16.42 a year.
 
-This paper looks at data from a 12 year period to determine the current 
-output of a specified city. It can be used to determine 
-on a month by month basis if a location produces a sufficient amount 
-of current from a solar panel. Providing customers with an idea on 
-whether or not they should make the investment that is solar.
+This paper looks at data from a 12 year period to determine the power 
+output of a small grid of solar panels
+located in Los Angeles, California,
+based on solar radiation data.
+The information gleaned from simulating power output
+can determine if,
+on an overall basis, a location produces a sufficient amount 
+of power from a solar panel.
+This gives customers information to determine
+whether or not they should make an investment in solar energy.
 
@@
 
-To translate raw irradiance data into meaningful current data, several steps
+To translate raw irradiance data into meaningful power data, several steps
 had to be taken to simulate the electrical workings of a real PV module. First,
@@
 df_high_irradiance <- df[df$Hour %in% HOUR_WINDOW, ]
-filtered_current_hist <- ggplot() +
-  geom_histogram(aes(x = df_high_irradiance$current_out, y = ..density..), binwidth = .0025) +
+filtered_power_hist <- ggplot() +
+  geom_histogram(aes(x = df_high_irradiance$power_out, y = ..density..), binwidth = .0025) +
   theme_bw() +
   labs(
-    title = "Distribution of Current Outputs", subtitle = "Restricted to high irradiance hours of the day",
-    x = "Current Output", y = "Density"
+    title = "Distribution of Power Outputs", subtitle = "Restricted to high irradiance hours of the day",
+    x = "Power Output", y = "Density"
   )
-print(filtered_current_hist)
+print(filtered_power_hist)
 ```
@@
 The main goal of this paper is to provide point and interval
-estimates for average (simulated) current output
+estimates for average (simulated) power output
 based on solar irradiance.
@@
 standard bootstrapping methods and dependent bootstrapping methods.
-The current output provides a good test case for this.
-Both the preprocessed and postprocessed currents are not independent data;
-values of either irradiance or current output
+The power output provides a good test case for this.
+Both the preprocessed GHI
+and postprocessed power outputs are not independent data;
+values of either irradiance or power output
 depend on surrounding values in time.
@@
 and $\theta$ is a statistic of interest.
-For our purposes, $\theta$ is simply the average current output
+For our purposes, $\theta$ is simply the average power output
 for our specific solar module configuration.
@@
 mostly independent blocks from which to sample:
-the daytime current values for each day.
-Given that our daytime current values for each day
+the daytime power values for each day.
+Given that our daytime power values for each day
 could be used as non-overlapping blocks,
@@
 time_series_ghi <- time_series_of_var(df, "GHI", "GHI", 6)
-time_series_current <- time_series_of_var(df, "current_out", "Output current", 6)
+time_series_power <- time_series_of_var(df, "power_out", "Output power", 6)
 ggsave("./images/time_series_ghi.png", make_plot_for_pres(time_series_ghi))
-ggsave("./images/time_series_current.png", make_plot_for_pres(time_series_current))
+ggsave("./images/time_series_power.png", make_plot_for_pres(time_series_power))
 ```
@@
 hour_plot_ghi <- hour_plot_of_var(df, "GHI", "GHI")
-hour_plot_current <- hour_plot_of_var(df, "current_out", "output current")
+hour_plot_power <- hour_plot_of_var(df, "power_out", "Output Power")
 ggsave("./images/hour_plot_ghi.png", make_plot_for_pres(hour_plot_ghi))
-ggsave("./images/hour_plot_current.png", make_plot_for_pres(hour_plot_current))
+ggsave("./images/hour_plot_power.png", make_plot_for_pres(hour_plot_power))
 ```
@@
 
-```{r current-ts, fig.cap="Time series of simulated current values for the first six days in the dataset.", echo=FALSE}
-print(time_series_current)
+```{r power-ts, fig.cap="Time series of simulated power values for the first six days in the dataset.", echo=FALSE}
+print(time_series_power)
 ```
@@
 
-```{r current-hour-boxplot, fig.cap="Distribution of GHI at each hour for the entire unfiltered dataset.", echo=FALSE}
-print(hour_plot_current)
+```{r power-hour-boxplot, fig.cap="Distribution of GHI at each hour for the entire unfiltered dataset.", echo=FALSE}
+print(hour_plot_power)
 ```
@@
     # blocks[i, ] <- df$current_out[(i - 1) * length_of_block + 1 : (i * length_of_block)]
-    blocks[i, ] <- df$current_out[start:stop]
+    blocks[i, ] <- df$power_out[start:stop]
     start <- start + length_of_block
@@
 iid_bootstrap_estimates_from_df <- function(df, B = 10000) {
-  boot_current <- boot(df$current_out, meanv, B)
-  return(boot_current$t)
+  boot_power <- boot(df$power_out, meanv, B)
+  return(boot_power$t)
 }
@@
   geom_histogram(aes(x = boot, y = ..density.., fill = type),
-    position = "identity", alpha = .5, color = "black",
-    binwidth = .0001
+    position = "identity", alpha = .5, color = "black"
   ) +
@@
   labs(
-    fill = "Bootstrap Type", title = "Bootstrapped Current Output", subtitle = "Percentile confidence intervals given",
+    fill = "Bootstrap Type", title = "Bootstrapped Power Output", subtitle = "Percentile confidence intervals given",
     x = "Bootstrap Sample Means", y = "Density"
@@
 
-```{r mean-boot-plot, fig.cap="Distributions of bootstrap mean simulated current for both iid and block bootstrapping methods.
-Percentile 95 percent bootstrap confidence intervals for mean simulated current overlayed.", echo=FALSE}
+```{r mean-boot-plot, fig.cap="Distributions of bootstrap mean simulated power for both iid and block bootstrapping methods.
+Percentile 95 percent bootstrap confidence intervals for mean simulated power overlayed.", echo=FALSE}
 print(mean_boot_plot)
@@
 bootstrap_ci_table %>% kable(booktabs = TRUE,
-caption="95% confidence intervals for the mean simulated current, as generated by each of the two bootstrapping methods.")
+caption="95% confidence intervals for the mean simulated power, as generated by each of the two bootstrapping methods.")
 
@@
 
-sixhr_plot_current <- time_series_irradiance(df_high_irradiance, "current_out", "Current Output", 6)
-ggsave("./images/sixhr_plot_current.png", make_plot_for_pres(sixhr_plot_current))
+sixhr_plot_power <- time_series_irradiance(df_high_irradiance, "power_out", "Power Output", 6)
+ggsave("./images/sixhr_plot_power.png", make_plot_for_pres(sixhr_plot_power))
 
@@
 df_more_irradiance <- df[df$Hour %in% HOUR_WINDOW, ]
-filtered_current_hist <- ggplot() +
-  geom_histogram(aes(x = df_high_irradiance$current_out)) +
+filtered_power_hist <- ggplot() +
+  geom_histogram(aes(x = df_high_irradiance$power_out)) +
   theme_bw()
-print(filtered_current_hist)
+print(filtered_power_hist)
 
-twelvehr_plot_current <- time_series_irradiance(df_more_irradiance, "current_out", "Current Output", 12)
-ggsave("./images/twelvehr_plot_current.png", make_plot_for_pres(twelvehr_plot_current))
+twelvehr_plot_power <- time_series_irradiance(df_more_irradiance, "power_out", "Power Output", 12)
+ggsave("./images/twelvehr_plot_power.png", make_plot_for_pres(twelvehr_plot_power))
 
-day_plot_current <- time_series_irradiance(df, "current_out", "Current Output", 24)
-ggsave("./images/day_plot_current.png", make_plot_for_pres(day_plot_current))
+day_plot_power <- time_series_irradiance(df, "power_out", "Power Output", 24)
+ggsave("./images/day_plot_power.png", make_plot_for_pres(day_plot_power))
 
@@
 However, if the results could be generalized,
-it would be good to see how current output based on solar irradiance
+it would be good to see how power output based on solar irradiance
 fits into a larger model of uncertainty surrounding renewable energy generation.

### 2021-12-16 deliverables/paper.Rmd
Rough discsussion

--- 2021-12-15 deliverables/paper.Rmd
@@
 
+# Discussion
+
+The power estimate generated in this study has practical applications in real-world electrical engineering problems.
+An allocation must be made that balances the expected contribution of renewable energy sources with the expected residual demand on the electrical grid.
+An intelligent estimate of the contribution from PV modules must account for geographical deviations in solar irradiance.
+While the estimate here was made for Los Angeles, this framework can be applied to any location for which irradiance data is tracked.
+This framework could also be applied to intervals of time where power estimates could be made for each hour of the day or for each day of the year; greater stratified specificity would improve the real-world usefulness of this method.
+Further work...
+
+I THINK DISCUSSION AND FUTURE WORK SHOULD BE COMBINED. THEY FLOW INTO EACH OTHER QUITE NICELY. FEEL FREE TO MERGE THEM IF YOU WANT.
+
 # Future Work

### 2021-12-16 deliverables/paper.Rmd
Added more to results section

@@
 
-# Preprocessing
+# Methods
+
+## Preprocessing
 
@@
 
-# Methods
-
 ## Data filtering
-
-```{r filter, include=FALSE}
-HOUR_WINDOW <- 11:16
-df_high_irradiance <- df[df$Hour %in% HOUR_WINDOW, ]
-filtered_power_hist <- ggplot() +
-  geom_histogram(aes(x = df_high_irradiance$power_out, y = ..density..), binwidth = .0025) +
-  theme_bw() +
-  labs(
-    title = "Distribution of Power Outputs", subtitle = "Restricted to high irradiance hours of the day",
-    x = "Power Output", y = "Density"
-  )
-print(filtered_power_hist)
-```
 
@@
 
+The data is obviously periodic.
+Both the GHI values and the simulated output power values
+were periodic by day.
+Both have a similar shape, 
+even with similar random variation,
+as is evidenced
+by Figure \ref{fig:ghi-ts} and Figure \ref{fig:power-ts}.
+This makes sense, as power output is a direct
+calculation (on our simulated model)
+from the solar radiance values.
+
 ```{r ghi-hour-boxplot, fig.cap="Distribution of GHI at each hour for the entire unfiltered dataset.", echo=FALSE}
@@
 print(hour_plot_power)
+```
+
+It should be noted that the hourly distribution
+of values for GHI and simulated power did not have
+exactly the same shape,
+as evidenced by Figure \ref{fig:ghi-hour-boxplot}
+and Figure \ref{power-hour-boxplot}.
+Both figures have values that,
+on average, move upward throughout the day,
+and then down towards zero at night.
+Nighttime accounts for the structural zeros seen the figures.
+The variation in the values over the day is due to expected
+variation in radiation due to the seasons,
+and random variation, like cloud cover.
+Some of the random variation can be seen in
+Figure \ref{fig:ghi-ts} and Figure \ref{fig:power-ts}.
+The range of distributions in Figure \ref{power-hour-boxplot}
+was used to guide the data filtering described in the next section.
+
+## Data Filtering
+
+```{r filter, include=FALSE}
+HOUR_WINDOW_LOW = 8
+HOUR_WINDOW_HIGH = 16
+HOUR_WINDOW <- HOUR_WINDOW_LOW:HOUR_WINDOW_HIGH
+
+df_high_irradiance <- df[df$Hour %in% HOUR_WINDOW, ]
+```
+
+Data was filtered to eliminate low- or no-irradiance hours
+from the dataset.
+The low-irradiance hours can be seen in
+Figure \ref{fig:unfiltered-hist}
+and Figure \ref{fig:ghi-hour-boxplot}.
+Filtering provides a more accurate measure of the power output
+during the daytime hours.
+The data was filtered to only include hours
+between `r HOUR_WINDOW_LOW` and `r HOUR_WINDOW_HIGH`,
+on a 24-hour timescale.
+The distribution of power output values after filtering
+can be seen in Figure \ref{fig:filtered-hist}.
+
+```{r unfiltered-hist, fig.cap = "Distribution of power outputs before filtering low-irradiance hours.", echo=FALSE}
+unfiltered_power_hist <- ggplot() +
+  geom_histogram(aes(x = df$power_out, y = ..density..)) +
+  theme_bw() +
+  labs(
+    x = "Power Output", y = "Density"
+  )
+print(unfiltered_power_hist)
+```
+
+```{r filtered-hist, fig.cap = "Distribution of power outputs after filtering low-irradiance hours.", echo=FALSE}
+filtered_power_hist <- ggplot() +
+  geom_histogram(aes(x = df_high_irradiance$power_out, y = ..density..)) +
+  theme_bw() +
+  labs(
+    x = "Power Output", y = "Density"
+  )
+print(filtered_power_hist)
 ```
@@
 }
+```
 
+```{r boot, include=FALSE}
 boot_block <- block_bootstrap_estimates_from_df(df_high_irradiance)
@@
 
+boot_mean_est <- mean(boot_block)
+```
+
+```{r vis, echo=FALSE}
 vis_df <- rbind(boot_block_df, boot_iid_df)
@@
 rownames(bootstrap_ci_table) <- c()
-
 bootstrap_ci_table %>% kable(booktabs = TRUE,
 caption="95% confidence intervals for the mean simulated power, as generated by each of the two bootstrapping methods.")
+```
 
+Bootstrapping was carried out with both the
+*iid* and block bootstrapping methods.
+95\% percentile confidence intervals for the mean power output
+based on both methods are shown in Table 1.
+It should be noted that the bounds for each method are
+similar to one another on an absolute scale,
+though on a relative scale of bootstrap estimates,
+the *iid* bootstrap produced a smaller interval.
+The interval widths for the bootstrap estimates,
+along with histograms of the bootstrap estimates,
+are shown in Figure \ref{fig:mean-boot-plot}.
+
+The bootstrapped average power value,
+based on the block bootstrap,
+was `r round(boot_mean_est, 3)` W.
+This provides an accurate estimate of the
+average power during daylight hours
+provided by the physical solar panel model.
+In particular, we report the estimated value
+from the block bootstrap because
+it accounts for time dependence.
+In this case the mean value between the two
+methods was very similar,
+as seen in Figure \ref{fig:mean-boot-plot}.
+However, given that we know this boostrapping
+method provides a better estimate for the density overall,
+it seems more sound to report the block bootstrap estimate.
+
+```{r power-output-est, echo=FALSE}
+power_out_est_day = boot_mean_est * length(HOUR_WINDOW)
+power_out_est_year = power_out_est_day * 365
 ```
+
+Based on our block bootstrap estimate,
+a solar panel in LA in our configuration
+would produce
+`r round(power_out_est_day)` Wh
+of energy over the course of a day.
+The panels would produce
+`r round(power_out_est_year)` Wh
+or 
+`r round(power_out_est_year/1000)` kWh
+of energy over the course of one year,
+on average.
 
@@
 
-HOUR_WINDOW <- 8:19
-df_more_irradiance <- df[df$Hour %in% HOUR_WINDOW, ]
-filtered_power_hist <- ggplot() +
-  geom_histogram(aes(x = df_high_irradiance$power_out)) +
-  theme_bw()
-print(filtered_power_hist)
-
-twelvehr_plot_power <- time_series_irradiance(df_more_irradiance, "power_out", "Power Output", 12)
+twelvehr_plot_power <- time_series_irradiance(df_high_irradiance, "power_out", "Power Output", 12)
 ggsave("./images/twelvehr_plot_power.png", make_plot_for_pres(twelvehr_plot_power))
@@
 ggsave("./images/day_plot_power.png", make_plot_for_pres(day_plot_power))
-
-# B1 <- df[df$Hour == c(7:10),]$current_output # Hours before peak
-# B2 <- df[df$Hour == c(11:14),]$current_output # Peak hours
-# B3 <- df[df$Hour == c(15:18),]$current_output # Hours after peak
-
-# nbb(hours, 4) # in order to make b blocks I would need to solve for b = [n/l] (floor) where n is the length of our data and l is 4, so make nbb call l instead of ib
-
-# nbb(hours, 12)
-
-# Could also block instead by hours but by total current output. Peak hours
-# would be a frame around the highest output and then hours before and after
-# would coincide with that. This would cause the hours to potentially be
-# different based on the season/month as months that get dark earlier could see
-# peak hours earlier in the day.
-
-# dpeak <- c(B1,B2,B3)
 ```

### 2021-12-16 deliverables/paper.Rmd
discussions
I ended up using just the HOUR_WINDOW as our range of peak hours.

@@
 
-The power estimate generated in this study has practical applications in real-world electrical engineering problems.
-An allocation must be made that balances the expected contribution of renewable energy sources with the expected residual demand on the electrical grid.
-An intelligent estimate of the contribution from PV modules must account for geographical deviations in solar irradiance.
-While the estimate here was made for Los Angeles, this framework can be applied to any location for which irradiance data is tracked.
-This framework could also be applied to intervals of time where power estimates could be made for each hour of the day or for each day of the year; greater stratified specificity would improve the real-world usefulness of this method.
-Further work...
+These estimates provide consumers with an answer to 
+a very pressing question. As discussed in the introduction, one of the biggest 
+problems about purchasing a solar panel is whether or not 
+it will end up paying for itself. One of the three stipulations in 
+order to calculate your average yearly return is a solar 
+panels output. After using bootstrapping we were able to 
+calculate a solar panel in Los Angeles' production 
+output being `r round(power_out_est_year/1000)`. Now, all 
+that is needed is the peak hours and total average cost/kWh. 
+The former can be calculated by setting a lower floor on 
+solar irradiance. Stated on Solar Reviews, 
+"1 peak sun hour" is equal to, "1000 W/m² of sunlight 
+per hour", meaning our length used initially for an hour window 
+would be considered peak hours. Taking this total of `r length(HOUR_WINDOW)` 
+peak hours and our average cost per kilowatt-hour of /$0.03 from before 
+we would find that on average a consumer in Los Angeles can 
+anticipate earning /$`r round(power_out_est_year/1000,2)*.03*length(HOUR_WINDOW)` yearly on one solar panel grid. 
+Knowing this, any resident located in or around the Los Angeles area should
+feel confident in their decision of purchasing a solar panel unit.
+
+The power estimate generated in this study has practical applications 
+in real-world electrical engineering problems. An allocation must be 
+made that balances the expected contribution of renewable energy 
+sources with the expected residual demand on the electrical grid. 
+An intelligent estimate of the contribution from PV modules must account 
+for geographical deviations in solar irradiance. While the estimate here 
+was made for Los Angeles, this framework can be applied to any location 
+for which irradiance data is tracked. This framework could also be applied to 
+intervals of time where power estimates could be made for each hour of the 
+day or for each day of the year; greater stratified specificity would 
+improve the real-world usefulness of this method.
+
+In regard to consumer usage, an automated website could be created 
+related to the information received in this paper. A simple 
+user input of location could pull all the values necessary to caluclate
+a predicied solar irradiance and transform that 
+information, as we did above, into an estimate of total yearly 
+solar panel energy cost return. This may not even just be important 
+information to jsut consumers, considering the push of renewable energy 
+research in the last few years by the US governement, this could be used
+to find the best areas in the US to began creating large-scale solar plants. 
+Finding the best city throughout the US in which NREL holds data on would turn 
+our problem into an optimization one.
 
@@
 ```
-
-- We chose our initial paper because of its use of Monte Carlo.
-After working with it for a little it was brought to our attention that Monte Carlo was used in conjunction with a backward/forward sweep method in order to calculate the pdf.
-After some confusion it was decided to switch to another paper dealing with the same data/motivation but without said methods (including Monte Carlo).
-- Luckily, the week after we switched papers we begun learning about bootstrapping, a way to sample from our data while taking into consideration our time change.
-
-INSERT GRAPHS FROM IID BOOTSTRAPPING AND BLOCK HERE
-ALSO INSERT FIG 1 FROM PAPER
-
-- The paper focused most of their sampling on night-day clustering, showing the difference between night and day through current.
-Night yielded thousands of 0 current values, leading to skewed data when sampling.
-Instead we took too approaches to bootstrap the current vs.
-hour data.
-- One using non-overlapping blocks sampling from the 8 hours surrounding the peak irradiance. The other being IID, yielding dependence in our data.

### 2021-12-16 deliverables/paper.Rmd
added disclaimer

@@
 "1 peak sun hour" is equal to, "1000 W/m² of sunlight 
-per hour", meaning our length used initially for an hour window 
-would be considered peak hours. Taking this total of `r length(HOUR_WINDOW)` 
+per hour" (@energysage). In our calculations we used an hour window between 
+8am to 4pm, a typical day range where the sun is overhead. A disclaimer,
+some of the values at the beginning and end of the day may not hold 
+exactly greater than our cutoff stated by Solar Reviews, however this range 
+yielded the greatest results of solar irradiance and Wh, some being much larger
+than 1000 W/m^2 and others being just under. Therefore, we are deciding to use 
+this range as our peak hours. Taking this total of `r length(HOUR_WINDOW)` 
 peak hours and our average cost per kilowatt-hour of /$0.03 from before 
@@
 Knowing this, any resident located in or around the Los Angeles area should
-feel confident in their decision of purchasing a solar panel unit.
+feel confident in their decision of purchasing a solar panel grid.
 

### 2021-12-16 deliverables/paper.Rmd
Some bug/formatting fixes

@@
 
-```{r unfiltered-hist, fig.cap = "Distribution of power outputs before filtering low-irradiance hours.", echo=FALSE}
+```{r unfiltered-hist, fig.cap = "Distribution of power outputs before filtering low-irradiance hours.", echo=FALSE, warning=FALSE}
 unfiltered_power_hist <- ggplot() +
@@
 
-```{r filtered-hist, fig.cap = "Distribution of power outputs after filtering low-irradiance hours.", echo=FALSE}
+```{r filtered-hist, fig.cap = "Distribution of power outputs after filtering low-irradiance hours.", echo=FALSE, warning=FALSE}
 filtered_power_hist <- ggplot() +
@@
 
-```{r vis, echo=FALSE}
+```{r vis, echo=FALSE, warning=FALSE}
 vis_df <- rbind(boot_block_df, boot_iid_df)

### 2021-12-16 deliverables/paper.Rmd
Formatting fix
combine discussion and future

@@
 
-```{r knitr-opts}
+```{r knitr-opts, echo=FALSE}
 knitr::opts_chunk$set(fig.pos = "t", fig.align = 'center', out.width = "75%", out.extra = "")
@@
 
-I THINK DISCUSSION AND FUTURE WORK SHOULD BE COMBINED. THEY FLOW INTO EACH OTHER QUITE NICELY. FEEL FREE TO MERGE THEM IF YOU WANT.
-
-# Future Work
-
 For future work, it would be good to examine the wider implications

### 2021-12-16 deliverables/paper.Rmd
Minor fix

@@
 this range as our peak hours. Taking this total of `r length(HOUR_WINDOW)` 
-peak hours and our average cost per kilowatt-hour of /$0.03 from before 
+peak hours and our average cost per kilowatt-hour of \$0.03 from before 
 we would find that on average a consumer in Los Angeles can 
-anticipate earning /$`r round(power_out_est_year/1000,2)*.03*length(HOUR_WINDOW)` yearly on one solar panel grid. 
+anticipate earning \$`r round(round(power_out_est_year/1000,2)*.03*length(HOUR_WINDOW),2)` yearly on one solar panel grid. 
 Knowing this, any resident located in or around the Los Angeles area should

### 2021-12-16 deliverables/paper.Rmd
Fixed formatting and grammar

@@
 ```{r knitr-opts, echo=FALSE}
-knitr::opts_chunk$set(fig.pos = "t", fig.align = 'center', out.width = "75%", out.extra = "")
+knitr::opts_chunk$set(fig.pos = "t", fig.align = 'center', out.width = "75%", out.extra = "",
+message=FALSE, warning=FALSE)
 ```
@@
 energy bill, and eventually have it pay for itself. In order to 
-calculate a solar panels net output experts recommend looking at three things. 
+calculate a solar panels net output experts recommend looking at three things:
+
 1. Determine your locations peak sun hours (average US around 4 hours/day)
@@
 3. Determine your solar panels net energy output
+
 Once you have calculated these three numbers it is as easy as 
@@
 as evidenced by Figure \ref{fig:ghi-hour-boxplot}
-and Figure \ref{power-hour-boxplot}.
+and Figure \ref{fig:power-hour-boxplot}.
 Both figures have values that,
@@
 Figure \ref{fig:ghi-ts} and Figure \ref{fig:power-ts}.
-The range of distributions in Figure \ref{power-hour-boxplot}
+The range of distributions in Figure \ref{fig:power-hour-boxplot}
 was used to guide the data filtering described in the next section.
@@
 unfiltered_power_hist <- ggplot() +
-  geom_histogram(aes(x = df$power_out, y = ..density..)) +
+  geom_histogram(aes(x = df$power_out, y = ..density..), bins=50) +
   theme_bw() +
@@
 filtered_power_hist <- ggplot() +
-  geom_histogram(aes(x = df_high_irradiance$power_out, y = ..density..)) +
+  geom_histogram(aes(x = df_high_irradiance$power_out, y = ..density..), bins=50) +
   theme_bw() +
@@
 
-```{r vis, echo=FALSE, warning=FALSE}
+```{r vis, echo=FALSE, warning=FALSE, message=FALSE}
 vis_df <- rbind(boot_block_df, boot_iid_df)
@@
 )
-colnames(bootstrap_ci_table) <- c("Bootstrap Method", "95% CI")
+colnames(bootstrap_ci_table) <- c("Bootstrap Method", "95 Percent CI")
 rownames(bootstrap_ci_table) <- c()
@@
 
+\pagebreak
+
 # Discussion
@@
 output being `r round(power_out_est_year/1000)`. Now, all 
-that is needed is the peak hours and total average cost/kWh. 
+that is needed is the peak hours and total average cost per kWh. 
 The former can be calculated by setting a lower floor on 
-solar irradiance. Stated on Solar Reviews, 
+solar irradiance. As stated on Solar Reviews, 
 "1 peak sun hour" is equal to, "1000 W/m² of sunlight 
 per hour" (@energysage). In our calculations we used an hour window between 
-8am to 4pm, a typical day range where the sun is overhead. A disclaimer,
+8am to 4pm, a typical day range where the sun is overhead. As a disclaimer,
 some of the values at the beginning and end of the day may not hold 
@@
 yielded the greatest results of solar irradiance and Wh, some being much larger
-than 1000 W/m^2 and others being just under. Therefore, we are deciding to use 
+than 1000 W/m^2 and others being just under. Therefore, we decided to use 
 this range as our peak hours. Taking this total of `r length(HOUR_WINDOW)` 
@@
 
-\newpage
+\pagebreak
 

### 2021-12-17 deliverables/paper.Rmd
Fixed grammar errors, added detail

@@
 utilizing the process of Monte Carlo sampling or other 
-statistcal computing methods. We landed on a few papers 
-that did exatcly this, one titled, "SolarStat: Modeling 
-Photovoltaic Sources through Stochastic Markov Processes". 
-Instead of comparing our works of Monte Carlo simulations 
-we were able to utilize a more recent topic discussed in class.
-This of course was bootsrapping, a topic we didn't get too in
-depth about; however, everything needed to know will be explained 
-later on in the paper. Regarding our choice of articles, it was a
-question of interest and importance. During the last decade the world
-has seen a massive uptick in renewable energy pushes. With this 
-new wave of renewables and the rather large misunderstanding of their
-total power we decided it would be interesting to look into solar and 
-what it was all about. The big question being asked regarding
+statistical computing methods. We landed on a few papers 
+that did exactly this, titled "SolarStat: Modeling 
+Photovoltaic Sources through Stochastic Markov Processes"
+(@miozzo_2014). 
+The paper's objectives were to simulate power output from
+a small grid of solar panels in Los Angeles using solar radiance data,
+and then cluster those results into different levels
+of radiance using Markov models.
+We adapted this approach.
+The goal of this paper was to use the power output
+simulation model from @miozzo_2014 to obtain similar
+power output results,
+and then to apply and compare two types of bootstrapping estimates,
+resulting in an overall estimate of power output.
+The two types of bootstrapping estimates
+are an *iid* bootstrap, which assumes the data is independent,
+and a block bootstrap, which assumes dependence in the data.
+We wanted to see in what ways these methods produced differing
+results, and hopefully obtain an accurate estimate of energy output.
+
+This problem is relevant to current trends in energy:
+During the last decade the world
+has seen a massive uptick in renewable energy pushes. 
+The big question being asked regarding
 solar power is whether or not it is worth heading the initial cost
-of set up and general mantienance. Many websites inform users of 
+of set up and general maintenance. Many websites inform users of 
 just this, according to the Solar.com Learning Center, the average
@@
 To translate raw irradiance data into meaningful power data, several steps
-had to be taken to simulate the electrical workings of a real PV module. First,
+had to be taken to simulate the electrical workings of a real PV (photovoltaic)
+module. First,
 adjustments had to be made to account for the fact that an imperfect amount of
@@
 
-## Effective Irradiance
+Note that, for all physical parameters given,
+@miozzo_2014 relied on the physical characteristics
+of a Solarbotics SCC-3733 Monocrystalline micro-solar panel
+[@miozzo_2014, p. 691].
+This means that all power output results should be
+weighed against this model panel.
+
+### Effective Irradiance
 
@@
 
-## PV Module
+### PV Module
 
@@
 ## Data filtering
+
+We assumed that based on the nature of solar data,
+the data would contain almost 50\% structural zeros,
+corresponding to nighttime hours.
+We decided that, after transforming GHI to power output,
+we would filter out the structural zeros
+based on the distribution of the output data,
+before analysis.
+
+This would provide a more accurate assessment of power
+output than simply averaging over a 24-hour period,
+in which many of the hours would have zero power output
+by nature of the day-night cycle.
 
@@
 anticipate earning \$`r round(round(power_out_est_year/1000,2)*.03*length(HOUR_WINDOW),2)` yearly on one solar panel grid. 
-Knowing this, any resident located in or around the Los Angeles area should
-feel confident in their decision of purchasing a solar panel grid.
+It is unclear how these results would scale with a solar panel
+or solar grid using different components.
+Our simulation was based on physical parameters provided
+in @miozzo_2014,
+which in turn only estimate the characteristics of
+the model solar panel.
+Furthermore, the model solar panel was a "micro-solar" panel:
+the 6x6 grid of panels had a reported size of
+26.96 cm$^2$, according to @miozzo_2014 [@miozzo_2014, p. 695].
+A larger or more efficient panel may be able to generate more
+total power,
+meaning that it may be possible for a consumer to purchase
+a larger solar panel with more output.
 
 The power estimate generated in this study has practical applications 
-in real-world electrical engineering problems. An allocation must be 
-made that balances the expected contribution of renewable energy 
-sources with the expected residual demand on the electrical grid. 
+in real-world electrical engineering problems.
+In real-world power grids,
+the expected contribution of renewable energy 
+sources 
+is balanced
+with the expected residual demand on the electrical grid. 
 An intelligent estimate of the contribution from PV modules must account 
@@
 related to the information received in this paper. A simple 
-user input of location could pull all the values necessary to caluclate
-a predicied solar irradiance and transform that 
+user input of location could pull all the values necessary to calculate
+a predicted solar irradiance and transform that 
 information, as we did above, into an estimate of total yearly 
 solar panel energy cost return. This may not even just be important 
-information to jsut consumers, considering the push of renewable energy 
-research in the last few years by the US governement, this could be used
+information to just consumers, considering the push of renewable energy 
+research in the last few years by the US government, this could be used
 to find the best areas in the US to began creating large-scale solar plants. 
-Finding the best city throughout the US in which NREL holds data on would turn 
-our problem into an optimization one.
+Finding the best city throughout the US in which NREL collects data 
+would allow power output to be optimized by location,
+as it likely is in practice.
 
@@
 of the inferential results explored in this paper.
-The model paper for this project simulated results for small solar panels
-in a specific configuration,
-making it hard to generalize the results to any meaningful solar array.
+As mentioned,
+the model paper for this project, @miozzo_2014, simulated
+results for small solar panels
+in a specific configuration.
+This makes it hard to generalize the results to any meaningful solar array.
 However, if the results could be generalized,
-it would be good to see how power output based on solar irradiance
+it would be useful to see how power output based on solar irradiance
 fits into a larger model of uncertainty surrounding renewable energy generation.

### 2021-12-17 deliverables/paper.Rmd
Added units and clarification

@@
 with the caveat that the calculations may only be approximate.
+GHI is measured in units of $W/m^2$, that is,
+power per unit area provided by solar radiation.
 
@@
 
-time_series_ghi <- time_series_of_var(df, "GHI", "GHI", 6)
-time_series_power <- time_series_of_var(df, "power_out", "Output power", 6)
+time_series_ghi <- time_series_of_var(df, "GHI", "GHI (W/m^2)", 6)
+time_series_power <- time_series_of_var(df, "power_out", "Output power (W)", 6)
 ggsave("./images/time_series_ghi.png", make_plot_for_pres(time_series_ghi))
@@
 
-hour_plot_ghi <- hour_plot_of_var(df, "GHI", "GHI")
-hour_plot_power <- hour_plot_of_var(df, "power_out", "Output Power")
+hour_plot_ghi <- hour_plot_of_var(df, "GHI", "GHI (W/m^2)")
+hour_plot_power <- hour_plot_of_var(df, "power_out", "Output Power (W)")
 ggsave("./images/hour_plot_ghi.png", make_plot_for_pres(hour_plot_ghi))
@@
   labs(
-    x = "Power Output", y = "Density"
+    x = "Power Output (W)", y = "Density"
   )
@@
 
-sixhr_plot_power <- time_series_irradiance(df_high_irradiance, "power_out", "Power Output", 6)
+sixhr_plot_power <- time_series_irradiance(df_high_irradiance, "power_out", "Power Output (W)", 6)
 ggsave("./images/sixhr_plot_power.png", make_plot_for_pres(sixhr_plot_power))
 
-twelvehr_plot_power <- time_series_irradiance(df_high_irradiance, "power_out", "Power Output", 12)
+twelvehr_plot_power <- time_series_irradiance(df_high_irradiance, "power_out", "Power Output (W)", 12)
 ggsave("./images/twelvehr_plot_power.png", make_plot_for_pres(twelvehr_plot_power))
 
-day_plot_power <- time_series_irradiance(df, "power_out", "Power Output", 24)
+day_plot_power <- time_series_irradiance(df, "power_out", "Power Output (W)", 24)
 ggsave("./images/day_plot_power.png", make_plot_for_pres(day_plot_power))
